---
title: "el_bias"
author: "Sophie Yam"
date: "`r Sys.Date()`"
output:
  github_document:
    html_preview: false
  pdf_document:
    number_sections: true
    toc: true
    toc_depth: 2
fontsize: 11pt
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
echo = TRUE,
message = FALSE,
warning = FALSE
)
```

# Libraries

```{r}
#install.packages("tabulapdf")

# Load required packages
library(tidyverse)
library(dplyr)
library(purrr)
library(lubridate)
library(ggplot2)
library(here)
library(parallel)
library(pdftools)
library(stringr)
library(rJava)
library(tabulapdf) # for data tables extraction
library(readxl)
```

# Data Extraction

Data for RRFs are coming from the US FDA Clap List.
<https://cdrh-rst.fda.gov/chemicals-list-analytical-performance-clap>

Each file is in a pdf format and contains RRF values for 3 concentration
of a surrogate. LCMS RRFs CLAP Negative Ion Detection.pdf (255.54 KB)
LCMS RRFs CLAP Positive Ion Detection.pdf (279.48 KB) RRF GCMS for
CLAP.pdf (322.77 KB)

They are first extracted to perform the bootstrap simulation. Some data
are n.d. or below DL. The current strategy is to change the n.d. to 0
and below DL to 0.001 based on the input of the Subject Matter Expert.

# Extract GC/MS data

Tried to extract the data directly from the pdf but it was complicated.
I manually performed the extraction in Excel using PowerQuery.

```{r}
gc_ms <- extract_tables("dataset/GCMS.pdf", method="stream")
gc_ms
```

# GC/MS data

```{r}
gcms_clap <- read_excel( 
  path = "processed_data/GCMS.xlsx", 
  sheet = "GC-MS")

gcms_clap_clean <- gcms_clap %>%
  na.omit()

gcms_clap_clean

gcms_clap_clean$"5 µg/mL" <- as.numeric(gcms_clap_clean$"5 µg/mL")
gcms_clap_clean$"10 µg/mL" <- as.numeric(gcms_clap_clean$"10 µg/mL")
gcms_clap_clean$"20 µg/mL" <- as.numeric(gcms_clap_clean$"20 µg/mL")

hist(gcms_clap_clean$"5 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 5 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "skyblue")
hist(gcms_clap_clean$"10 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 10 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "orange")
hist(gcms_clap_clean$"20 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 20 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "red")


```

# LC/MS pos data

```{r}
lcmspos_clap <- read_excel( 
  path = "processed_data/LCMS.xlsx", 
  sheet = "LCMSPOS")

lcmspos_clap_clean <- lcmspos_clap %>%
  na.omit()


hist(lcmspos_clap_clean$"5 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 5 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "skyblue")
hist(lcmspos_clap_clean$"10 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 10 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "orange")
hist(lcmspos_clap_clean$"20 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 20 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "red")

```

# LC/MS neg data

```{r}
lcmsneg_clap <- read_excel( 
  path = "processed_data/LCMS.xlsx", 
  sheet = "LCMSNEG")

lcmsneg_clap_clean <- lcmsneg_clap %>%
  na.omit()

lcmsneg_clap_clean


hist(lcmsneg_clap_clean$"5 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 5 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "skyblue")
hist(lcmsneg_clap_clean$"10 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 10 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "orange")
hist(lcmsneg_clap_clean$"20 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 20 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "red")

```

# Generate Base concentration

## Define AET

```{r}
AET <- 5
```


## Uniform distribution

```{r}
set.seed(123)
base_concentrations1 <- runif(50, min = 0.5 * AET, max = 2 * AET)
hist(base_concentrations1)
```

## Log-Normal distribution

$log(0.001) = -6.907, log(100) = 4.605$ \\ 
$mid (meanlog) approx -1.15$ \\
$width = 4.605 - (-6.907) = 11.51$ \\
$width/6 = 11.51/6=1.91$

```{r}
set.seed(123)
base_concentrations2 <- rlnorm(50, meanlog = -1.15, sdlog = 1.91)
hist(base_concentrations2)
```

## Custom distribution

```{r}
set.seed(123)
Z <- runif(70, min = 0.001, max = 1)
Y <- runif(20, min = 1, max = 10)
K <- runif(10, min = 10, max = 100)

custom <- c(Z, Y, K)

custom_sampled <- sample(custom, size = 100, replace = FALSE)

hist(custom_sampled)
```

## Normal distribution (simulating a GC/MS RRF distribution)

```{r}
set.seed(123)
RRF_norm <- rnorm(50, 1, 0.2)
RRF_norm[RRF_norm<0] <- 0.001 # avoiding negative numbers
hist(RRF_norm)
```

## Log_Normal distribution (simulating a LC/MS RRF distribution)

```{r}
set.seed(123)
RRF_lnorm <- rlnorm(50, 0, 0.5)
RRF_lnorm[RRF_lnorm<0] <- 0.001 # avoiding negative numbers
hist(RRF_lnorm)
```

# Scenario with uniform distribution & GC/MS data from US FDA CLAP list


## Initialisation of the comparison table

```{r}
summary_table <- data.frame(
  Method = character(),
  FN_Rate = numeric(),
  FP_Rate = numeric(),
  FN_SD = numeric(),
  FP_SD = numeric()
)
```

## GC-MS No Correction Scenarios

Uniform concentration around the AET \\ 
Non Parametric Bootstrap with replacement of the GC/MS dataset \\
No UF correction

```{r}
set.seed(123)
N_sim <- 10000

## 1) Concentration around the AET
base_conc_sim <- sample(base_concentrations1, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## alpha) No Correction
corrected_conc <- base_conc_sim * simulated_RRF

FN <- (base_conc_sim > AET) & (corrected_conc < AET)
FP <- (base_conc_sim< AET) & (corrected_conc > AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "No Correction",
  FN_Rate = mean(FN) * 100,
  FP_Rate = mean(FP) * 100,
  FN_SD = (sd(FN)/sqrt(N_sim))*100,
  FP_SD = (sd(FP)/sqrt(N_sim))*100
))

```

```{r}
hist(base_conc_sim, 
     main = "Histogram of Base concentration around the AET", 
     xlab = "Concentration",
     col = "skyblue")
hist(simulated_RRF,
     main = "Histogram of RRFs (non-parametric bootstrap of GC/MS dataset 5µg/mL) from FDA Clap list", 
     xlab = "Concentration",
     col = "grey")
```

## GC-MS Correction UF Factor

Uniform concentration around the AET \\
Non Parametric Bootstrap with replacement of the GC/MS dataset \\
UF correction : from 2, 4, 5, 10 and UF = 1/(1-RSD)

```{r}
rsd <- function(x) {
  return(sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE))
}

set.seed(123)
N_sim <- 10000
UF_RSD <- 1/(1-rsd(simulated_RRF)) # formula from ISO 10993-18
UF <- c(2, 4, 5, 10, UF_RSD)

## 1) Concentration around the AET
base_conc_sim <- sample(base_concentrations1, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## beta) UF corrections

results_df <- data.frame(UF = numeric(), FN_Rate = numeric(), FP_Rate = numeric())

for (uf_val in UF) {
  corrected_conc_uf <- (base_conc_sim * simulated_RRF) * uf_val
  
  FN_loop <- (base_conc_sim > AET) & (corrected_conc_uf < AET)
  FP_loop <- (base_conc_sim < AET) & (corrected_conc_uf > AET)
  
  method_name <- ifelse(uf_val == UF_RSD, "UF (RSD)", paste0("UF (", round(uf_val, 1), ")"))

  summary_table <- rbind(summary_table, data.frame(
    Method = method_name,
    FN_Rate = mean(FN_loop) * 100,
    FP_Rate = mean(FP_loop) * 100,
    FN_SD = (sd(FN_loop)/sqrt(N_sim))*100,
    FP_SD = (sd(FP_loop)/sqrt(N_sim))*100
  ))

}



UF_RSD
```

## GC-MS Correction Factor percentile

Uniform concentration around the AET \\ 
Non Parametric Bootstrap with replacement of the GC/MS dataset \\ 
Correction by dividing 16th percentile

```{r}
percentile <- function(x) {
  return(quantile(x, .16))
}

set.seed(123)
N_sim <- 10000

## 1) Concentration around the AET
base_conc_sim <- sample(base_concentrations1, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## gamma) 16th percentile Correction
corrected_conc <- base_conc_sim * simulated_RRF /  percentile(simulated_RRF)

FN <- (base_conc_sim > AET) & (corrected_conc < AET)
FP <- (base_conc_sim < AET) & (corrected_conc > AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "Percentile (16th)",
  FN_Rate = mean(FN) * 100,
  FP_Rate = mean(FP) * 100,
  FN_SD = (sd(FN)/sqrt(N_sim))*100,
  FP_SD = (sd(FP)/sqrt(N_sim))*100
))

```

## GC-MS Correction RRFlow

Uniform concentration around the AET \\
Non Parametric Bootstrap with replacement of the GC/MS dataset \\
Correction by applying condition on the RRFs.

```{r}
set.seed(123)
N_sim <- 10000

## 1) Concentration around the AET
base_conc_sim <- sample(base_concentrations1, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## delta) RRflow condition

corrected_conc <- base_conc_sim
condition <- (simulated_RRF < 0.5) | (simulated_RRF > 2)
corrected_conc[condition] <- corrected_conc[condition] / simulated_RRF[condition]


FN <- (base_conc_sim > AET) & (corrected_conc < AET)
FP <- (base_conc_sim < AET) & (corrected_conc > AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "RRFlow",
  FN_Rate = mean(FN) * 100,
  FP_Rate = mean(FP) * 100,
  FN_SD = (sd(FN)/sqrt(N_sim))*100,
  FP_SD = (sd(FP)/sqrt(N_sim))*100
))

```
## GC-MS Correction RRF < 1

Uniform concentration around the AET \\
Non Parametric Bootstrap with replacement of the GC/MS dataset \\
Correction by applying condition on the RRFs.

```{r}
set.seed(123)
N_sim <- 10000

## 1) Concentration around the AET
base_conc_sim <- sample(base_concentrations1, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## epsilon RRF < 1

corrected_conc <- base_conc_sim
condition <- (simulated_RRF < 1)
corrected_conc[condition] <- corrected_conc[condition] / simulated_RRF[condition]


FN <- (base_conc_sim > AET) & (corrected_conc < AET)
FP <- (base_conc_sim < AET) & (corrected_conc > AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "RRF<1",
  FN_Rate = mean(FN) * 100,
  FP_Rate = mean(FP) * 100,
  FN_SD = (sd(FN)/sqrt(N_sim))*100,
  FP_SD = (sd(FP)/sqrt(N_sim))*100
))


summary_table
```

## Comparison

```{r}
plot_data <- summary_table %>%
  pivot_longer(cols = c(FN_Rate, FP_Rate), names_to = "Type", values_to = "Rate") %>%
  mutate(
    SD = ifelse(Type == "FN_Rate", round(FN_SD, 2), round(FP_SD, 2)),
    Type = ifelse(Type == "FN_Rate", "False Negative", "False Positive")
  )

method_levels <- c(
  "No Correction",
  "UF (2)", 
  "UF (4)", 
  "UF (5)", 
  "UF (10)",  
  "UF (RSD)",
  "Percentile (16th)", 
  "RRFlow",
  "RRF<1"
)

plot_data$Method <- factor(plot_data$Method, levels = method_levels)


ggplot(plot_data, aes(x = Method, y = Rate, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), alpha = 0.8) +
  geom_errorbar(aes(ymin = pmax(0, Rate - SD), ymax = Rate + SD), 
                position = position_dodge(width = 0.9), width = 0.2) +
  labs(
    title = "Comparison of GC-MS Correction Methods",
    subtitle = "Conc. Uniform distribution (around AET), RRF bootstrapped data from CLAP list",
    y = "Rate (%)",
    x = "Method") +
  geom_text(aes(label = sprintf("%.1f%%", Rate)), position = position_dodge(width = 0.9), hjust = -0.5, size = 3, vjust = 0.1) +
  scale_fill_manual(values = c("False Negative" = "red", "False Positive" = "grey")) + 
  scale_y_continuous(limits = c(0, 100)) +
  theme_minimal() +
  coord_flip()

```


# Scenario with Log-Normal distribution & GC/MS data from US FDA CLAP list


## Initialisation of the comparison table

```{r}
summary_table <- data.frame(
  Method = character(),
  FN_Rate = numeric(),
  FP_Rate = numeric(),
  FN_SD = numeric(),
  FP_SD = numeric()
)
```

## GC-MS No Correction Scenarios

Non Parametric Bootstrap with replacement of the GC/MS dataset \\
No UF correction

```{r}
set.seed(123)
N_sim <- 10000

## 2) Log_Normal Concentration
base_conc_sim <- sample(base_concentrations2, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## alpha) No Correction
corrected_conc <- base_conc_sim * simulated_RRF

FN <- (base_conc_sim > AET) & (corrected_conc < AET)
FP <- (base_conc_sim< AET) & (corrected_conc > AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "No Correction",
  FN_Rate = mean(FN) * 100,
  FP_Rate = mean(FP) * 100,
  FN_SD = (sd(FN)/sqrt(N_sim))*100,
  FP_SD = (sd(FP)/sqrt(N_sim))*100
))

```

```{r}
hist(base_conc_sim, 
     main = "Histogram of Base concentration around the AET", 
     xlab = "Concentration",
     col = "skyblue")
hist(simulated_RRF,
     main = "Histogram of RRFs (non-parametric bootstrap of GC/MS dataset 5µg/mL) from FDA Clap list", 
     xlab = "Concentration",
     col = "grey")
```

## GC-MS Correction UF Factor

Non Parametric Bootstrap with replacement of the GC/MS dataset \\
UF correction : from 2, 4, 5, 10 and UF = 1/(1-RSD)

```{r}
rsd <- function(x) {
  return(sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE))
}

set.seed(123)
N_sim <- 10000
UF_RSD <- 1/(1-rsd(simulated_RRF)) # formula from ISO 10993-18
UF <- c(2, 4, 5, 10, UF_RSD)

## 2) Log_Normal Concentration
base_conc_sim <- sample(base_concentrations2, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## beta) UF corrections

results_df <- data.frame(UF = numeric(), FN_Rate = numeric(), FP_Rate = numeric())

for (uf_val in UF) {
  corrected_conc_uf <- (base_conc_sim * simulated_RRF) * uf_val
  
  FN_loop <- (base_conc_sim > AET) & (corrected_conc_uf < AET)
  FP_loop <- (base_conc_sim < AET) & (corrected_conc_uf > AET)
  
  method_name <- ifelse(uf_val == UF_RSD, "UF (RSD)", paste0("UF (", round(uf_val, 1), ")"))

  summary_table <- rbind(summary_table, data.frame(
    Method = method_name,
    FN_Rate = mean(FN_loop) * 100,
    FP_Rate = mean(FP_loop) * 100,
    FN_SD = (sd(FN_loop)/sqrt(N_sim))*100,
    FP_SD = (sd(FP_loop)/sqrt(N_sim))*100
  ))

}



UF_RSD
```

## GC-MS Correction Factor percentile

Non Parametric Bootstrap with replacement of the GC/MS dataset \\ 
Correction by dividing 16th percentile

```{r}
percentile <- function(x) {
  return(quantile(x, .16))
}

set.seed(123)
N_sim <- 10000

## 2) Log_Normal Concentration
base_conc_sim <- sample(base_concentrations2, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## gamma) 16th percentile Correction
corrected_conc <- base_conc_sim * simulated_RRF /  percentile(simulated_RRF)

FN <- (base_conc_sim > AET) & (corrected_conc < AET)
FP <- (base_conc_sim < AET) & (corrected_conc > AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "Percentile (16th)",
  FN_Rate = mean(FN) * 100,
  FP_Rate = mean(FP) * 100,
  FN_SD = (sd(FN)/sqrt(N_sim))*100,
  FP_SD = (sd(FP)/sqrt(N_sim))*100
))

```

## GC-MS Correction RRFlow

Non Parametric Bootstrap with replacement of the GC/MS dataset \\
Correction by applying condition on the RRFs.

```{r}
set.seed(123)
N_sim <- 10000

## 2) Log_Normal Concentration
base_conc_sim <- sample(base_concentrations2, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## delta) RRflow condition

corrected_conc <- base_conc_sim
condition <- (simulated_RRF < 0.5) | (simulated_RRF > 2)
corrected_conc[condition] <- corrected_conc[condition] / simulated_RRF[condition]


FN <- (base_conc_sim > AET) & (corrected_conc < AET)
FP <- (base_conc_sim < AET) & (corrected_conc > AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "RRFlow",
  FN_Rate = mean(FN) * 100,
  FP_Rate = mean(FP) * 100,
  FN_SD = (sd(FN)/sqrt(N_sim))*100,
  FP_SD = (sd(FP)/sqrt(N_sim))*100
))

```

## GC-MS Correction RRF < 1

Non Parametric Bootstrap with replacement of the GC/MS dataset \\
Correction by applying condition on the RRFs.

```{r}
set.seed(123)
N_sim <- 10000

## 2) Log_Normal Concentration
base_conc_sim <- sample(base_concentrations2, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## epsilon) RRF<1

corrected_conc <- base_conc_sim
condition <- (simulated_RRF < 1)
corrected_conc[condition] <- corrected_conc[condition] / simulated_RRF[condition]


FN <- (base_conc_sim > AET) & (corrected_conc < AET)
FP <- (base_conc_sim < AET) & (corrected_conc > AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "RRF<1",
  FN_Rate = mean(FN) * 100,
  FP_Rate = mean(FP) * 100,
  FN_SD = (sd(FN)/sqrt(N_sim))*100,
  FP_SD = (sd(FP)/sqrt(N_sim))*100
))

```


## Comparison

```{r}
plot_data <- summary_table %>%
  pivot_longer(cols = c(FN_Rate, FP_Rate), names_to = "Type", values_to = "Rate") %>%
  mutate(
    SD = ifelse(Type == "FN_Rate", round(FN_SD, 2), round(FP_SD, 2)),
    Type = ifelse(Type == "FN_Rate", "False Negative", "False Positive")
  )

method_levels <- c(
  "No Correction",
  "UF (2)", 
  "UF (4)", 
  "UF (5)", 
  "UF (10)",  
  "UF (RSD)",
  "Percentile (16th)", 
  "RRFlow",
  "RRF<1"
)

plot_data$Method <- factor(plot_data$Method, levels = method_levels)


ggplot(plot_data, aes(x = Method, y = Rate, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), alpha = 0.8) +
  geom_errorbar(aes(ymin = pmax(0, Rate - SD), ymax = Rate + SD), 
                position = position_dodge(width = 0.9), width = 0.2) +
  labs(
    title = "Comparison of GC-MS Correction Methods",
    subtitle = "Conc. Log-Normal distribution, RRF bootstrapped data from CLAP list",
    y = "Rate (%)",
    x = "Method") +
  geom_text(aes(label = sprintf("%.1f%%", Rate)), position = position_dodge(width = 0.9), hjust = -0.5, size = 3, vjust = 0.1) +
  scale_fill_manual(values = c("False Negative" = "red", "False Positive" = "grey")) + 
  scale_y_continuous(limits = c(0, 100)) +
  theme_minimal() +
  coord_flip()

```


# Scenario with Custom Dist. & GC/MS data from US FDA CLAP list


## Initialisation of the comparison table

```{r}
summary_table <- data.frame(
  Method = character(),
  FN_Rate = numeric(),
  FP_Rate = numeric(),
  FN_SD = numeric(),
  FP_SD = numeric()
)
```

## GC-MS No Correction Scenarios

Non Parametric Bootstrap with replacement of the GC/MS dataset \\
No UF correction

```{r}
set.seed(123)
N_sim <- 10000

## 3) Concentration from custom dist.
base_conc_sim <- sample(custom_sampled, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## alpha) No Correction
corrected_conc <- base_conc_sim * simulated_RRF

FN <- (base_conc_sim > AET) & (corrected_conc < AET)
FP <- (base_conc_sim< AET) & (corrected_conc > AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "No Correction",
  FN_Rate = mean(FN) * 100,
  FP_Rate = mean(FP) * 100,
  FN_SD = (sd(FN)/sqrt(N_sim))*100,
  FP_SD = (sd(FP)/sqrt(N_sim))*100
))

```

```{r}
hist(base_conc_sim, 
     main = "Histogram of Base concentration around the AET", 
     xlab = "Concentration",
     col = "skyblue")
hist(simulated_RRF,
     main = "Histogram of RRFs (non-parametric bootstrap of GC/MS dataset 5µg/mL) from FDA Clap list", 
     xlab = "Concentration",
     col = "grey")
```

## GC-MS Correction UF Factor

Uniform concentration around the AET \\
Non Parametric Bootstrap with replacement of the GC/MS dataset \\
UF correction : from 2, 4, 5, 10 and UF = 1/(1-RSD)

```{r}
rsd <- function(x) {
  return(sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE))
}

set.seed(123)
N_sim <- 10000
UF_RSD <- 1/(1-rsd(simulated_RRF)) # formula from ISO 10993-18
UF <- c(2, 4, 5, 10, UF_RSD)

## 3) Concentration from custom dist.
base_conc_sim <- sample(custom_sampled, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## beta) UF corrections

results_df <- data.frame(UF = numeric(), FN_Rate = numeric(), FP_Rate = numeric())

for (uf_val in UF) {
  corrected_conc_uf <- (base_conc_sim * simulated_RRF) * uf_val
  
  FN_loop <- (base_conc_sim > AET) & (corrected_conc_uf < AET)
  FP_loop <- (base_conc_sim < AET) & (corrected_conc_uf > AET)
  
  method_name <- ifelse(uf_val == UF_RSD, "UF (RSD)", paste0("UF (", round(uf_val, 1), ")"))

  summary_table <- rbind(summary_table, data.frame(
    Method = method_name,
    FN_Rate = mean(FN_loop) * 100,
    FP_Rate = mean(FP_loop) * 100,
    FN_SD = (sd(FN_loop)/sqrt(N_sim))*100,
    FP_SD = (sd(FP_loop)/sqrt(N_sim))*100
  ))

}



UF_RSD
```

## GC-MS Correction Factor percentile

Uniform concentration around the AET \\ 
Non Parametric Bootstrap with replacement of the GC/MS dataset \\ 
Correction by dividing 16th percentile

```{r}
percentile <- function(x) {
  return(quantile(x, .16))
}

set.seed(123)
N_sim <- 10000

## 3) Concentration from custom dist.
base_conc_sim <- sample(custom_sampled, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## gamma) 16th percentile Correction
corrected_conc <- base_conc_sim * simulated_RRF /  percentile(simulated_RRF)

FN <- (base_conc_sim > AET) & (corrected_conc < AET)
FP <- (base_conc_sim < AET) & (corrected_conc > AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "Percentile (16th)",
  FN_Rate = mean(FN) * 100,
  FP_Rate = mean(FP) * 100,
  FN_SD = (sd(FN)/sqrt(N_sim))*100,
  FP_SD = (sd(FP)/sqrt(N_sim))*100
))

```

## GC-MS Correction RRFlow
Non Parametric Bootstrap with replacement of the GC/MS dataset \\
Correction by applying condition on the RRFs.

```{r}
set.seed(123)
N_sim <- 10000

## 3) Concentration from custom dist.
base_conc_sim <- sample(custom_sampled, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## delta) RRflow condition

corrected_conc <- base_conc_sim
condition <- (simulated_RRF < 0.5) | (simulated_RRF > 2)
corrected_conc[condition] <- corrected_conc[condition] / simulated_RRF[condition]


FN <- (base_conc_sim > AET) & (corrected_conc < AET)
FP <- (base_conc_sim < AET) & (corrected_conc > AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "RRFlow",
  FN_Rate = mean(FN) * 100,
  FP_Rate = mean(FP) * 100,
  FN_SD = (sd(FN)/sqrt(N_sim))*100,
  FP_SD = (sd(FP)/sqrt(N_sim))*100
))

```


## GC-MS Correction RRF < 1

```{r}
set.seed(123)
N_sim <- 10000

## 3) Concentration from custom dist.
base_conc_sim <- sample(custom_sampled, size = N_sim, replace = TRUE)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_sim, replace = TRUE)
## delta) Correction RRF < 1

corrected_conc <- base_conc_sim
condition <- (simulated_RRF < 1)
corrected_conc[condition] <- corrected_conc[condition] / simulated_RRF[condition]


FN <- (base_conc_sim > AET) & (corrected_conc < AET)
FP <- (base_conc_sim < AET) & (corrected_conc > AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "RRF<1",
  FN_Rate = mean(FN) * 100,
  FP_Rate = mean(FP) * 100,
  FN_SD = (sd(FN)/sqrt(N_sim))*100,
  FP_SD = (sd(FP)/sqrt(N_sim))*100
))

```


## Comparison

```{r}
plot_data <- summary_table %>%
  pivot_longer(cols = c(FN_Rate, FP_Rate), names_to = "Type", values_to = "Rate") %>%
  mutate(
    SD = ifelse(Type == "FN_Rate", round(FN_SD, 2), round(FP_SD, 2)),
    Type = ifelse(Type == "FN_Rate", "False Negative", "False Positive")
  )

method_levels <- c(
  "No Correction",
  "UF (2)", 
  "UF (4)", 
  "UF (5)", 
  "UF (10)",  
  "UF (RSD)",
  "Percentile (16th)", 
  "RRFlow",
  "RRF<1"
)

plot_data$Method <- factor(plot_data$Method, levels = method_levels)


ggplot(plot_data, aes(x = Method, y = Rate, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), alpha = 0.8) +
  geom_errorbar(aes(ymin = pmax(0, Rate - SD), ymax = Rate + SD), 
                position = position_dodge(width = 0.9), width = 0.2) +
  labs(
    title = "Comparison of GC-MS Correction Methods",
    subtitle = "Conc. custom distribution, RRF bootstrapped data from CLAP list",
    y = "Rate (%)",
    x = "Method") +
  geom_text(aes(label = sprintf("%.1f%%", Rate)), position = position_dodge(width = 0.9), hjust = -0.5, size = 3, vjust = 0.1) +
  scale_fill_manual(values = c("False Negative" = "red", "False Positive" = "grey")) + 
  scale_y_continuous(limits = c(0, 100)) +
  theme_minimal() +
  coord_flip()

```








