---
title: "el_bias"
format: gfm
editor: visual
---

# Libraries

```{r}
Sys.setenv(RETICULATE_PYTHON = "C:/Users/misak/anaconda3/python.exe")
#install.packages("tabulapdf")

# Load required packages
library(tidyverse)
library(dplyr)
library(purrr)
library(lubridate)
library(ggplot2)
library(here)
library(parallel)
library(pdftools)
library(stringr)
library(rJava)
library(tabulapdf) # for data tables extraction
library(readxl)
library(reticulate)
#py_install(c("pandas", "joblib", "scikit-learn"))

```

# Data Extraction

Data for RRFs are coming from the US FDA Clap List. <https://cdrh-rst.fda.gov/chemicals-list-analytical-performance-clap>

Each file is in a pdf format and contains RRF values for 3 concentration of a surrogate. LCMS RRFs CLAP Negative Ion Detection.pdf (255.54 KB) LCMS RRFs CLAP Positive Ion Detection.pdf (279.48 KB) RRF GCMS for CLAP.pdf (322.77 KB)

They are first extracted to perform the bootstrap simulation. Some data are n.d. or below DL. The current strategy is to change the n.d. to 0 and below DL to 0.001 based on the input of the Subject Matter Expert.

# GC/MS data

```{r}
gcms_clap <- read_excel( 
  path = "processed_data/GCMS.xlsx", 
  sheet = "GC-MS")

gcms_clap_clean <- gcms_clap %>%
  na.omit()

gcms_clap_clean

gcms_clap_clean$"5 µg/mL" <- as.numeric(gcms_clap_clean$"5 µg/mL")
gcms_clap_clean$"10 µg/mL" <- as.numeric(gcms_clap_clean$"10 µg/mL")
gcms_clap_clean$"20 µg/mL" <- as.numeric(gcms_clap_clean$"20 µg/mL")

hist(gcms_clap_clean$"5 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 5 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "skyblue")
hist(gcms_clap_clean$"10 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 10 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "orange")
hist(gcms_clap_clean$"20 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 20 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "red")

```

# LC/MS pos data

```{r}
lcmspos_clap <- read_excel( 
  path = "processed_data/LCMS.xlsx", 
  sheet = "LCMSPOS")

lcmspos_clap_clean <- lcmspos_clap %>%
  na.omit()


hist(lcmspos_clap_clean$"5 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 5 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "skyblue")
hist(lcmspos_clap_clean$"10 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 10 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "orange")
hist(lcmspos_clap_clean$"20 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 20 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "red")

```

# LC/MS neg data

```{r}
lcmsneg_clap <- read_excel( 
  path = "processed_data/LCMS.xlsx", 
  sheet = "LCMSNEG")

lcmsneg_clap_clean <- lcmsneg_clap %>%
  na.omit()

lcmsneg_clap_clean


hist(lcmsneg_clap_clean$"5 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 5 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "skyblue")
hist(lcmsneg_clap_clean$"10 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 10 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "orange")
hist(lcmsneg_clap_clean$"20 µg/mL", 
     main = "Histogram of RRFs for GC/MS with a 20 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "red")

```

# Generate Base concentration

## Define AET

```{r}
AET <- 1
```

## Uniform distribution

```{r}
set.seed(123)
base_concentrations1 <- runif(100, min = 0.5 * AET, max = 2 * AET)
hist(base_concentrations1)
```

## Log-Normal distribution

$log(0.001) = -6.907, log(100) = 4.605$ \\ $mid (meanlog) approx -1.15$ \\ $width = 4.605 - (-6.907) = 11.51$ \\ $width/6 = 11.51/6=1.91$

```{r}
set.seed(123)
base_concentrations2 <- rlnorm(100, meanlog = -1.15, sdlog = 1.91)
hist(base_concentrations2)
```

## Custom distribution

```{r}
set.seed(123)
Z <- runif(70, min = 0.001, max = 1)
Y <- runif(20, min = 1, max = 10)
K <- runif(10, min = 10, max = 100)

custom <- c(Z, Y, K)

custom_sampled <- sample(custom, size = 100, replace = FALSE)

hist(custom_sampled)
```

## Normal distribution (simulating a GC/MS RRF distribution)

```{r}
set.seed(123)
RRF_norm <- rnorm(50, 1, 0.2)
RRF_norm[RRF_norm<0] <- 0.001 # avoiding negative numbers
hist(RRF_norm)
```

## Log_Normal distribution (simulating a LC/MS RRF distribution)

```{r}
set.seed(123)
RRF_lnorm <- rlnorm(50, 0, 0.5)
RRF_lnorm[RRF_lnorm<0] <- 0.001 # avoiding negative numbers
hist(RRF_lnorm)
```

# Scenario with uniform distribution & GC/MS data from US FDA CLAP list

## Initialisation of the comparison table

```{r}
summary_table <- data.frame(
  Method = character(),
  FN_Rate = numeric(),
  FP_Rate = numeric(),
  FN_SD = numeric(),
  FP_SD = numeric(),
  precision = numeric(),
  recall = numeric(),
  F1 = numeric(),
  f_beta = numeric()
)
```

## GC-MS No Correction Scenarios

```{r}
set.seed(123)
N_sim <- 10000
N_extractables <- 80
total_FN <- 0
total_FP <- 0
total_TP <- 0
total_TN <- 0


for (i in 1:N_sim) {
## 2) Concentration from the log-normal distribution
base_conc_sim <- rlnorm(N_extractables, meanlog = -1.15, sdlog = 1.91)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_extractables, replace = FALSE)
## I) No Correction
corrected_conc <- base_conc_sim * simulated_RRF

is_above_threshold <- base_conc_sim > AET
reported_above_threshold <- corrected_conc > AET

  total_FN <- total_FN + sum(is_above_threshold & !reported_above_threshold, na.rm = TRUE)
  total_FP <- total_FP + sum(!is_above_threshold & reported_above_threshold, na.rm = TRUE)
  total_TP <- total_TP + sum(is_above_threshold & reported_above_threshold, na.rm = TRUE)
  total_TN <- total_TN + sum(!is_above_threshold & !reported_above_threshold, na.rm = TRUE)
}

count_Above_AET <- total_TP + total_FN
count_Below_AET  <- total_TN + total_FP
FN_Rate_final <- total_FN / count_Above_AET
FP_Rate_final <- total_FP / count_Below_AET

recall <- total_TP / count_Above_AET
precision <- total_TP / (total_TP + total_FP)

F1 <- (2 * precision * recall / (precision + recall))
beta <- 5
f_beta <- (1 + beta^2) * (precision * recall) / ((beta^2 * precision) + recall)

SD_FN <- sqrt(FN_Rate_final * (1 - FN_Rate_final) / count_Above_AET)
SD_FP <- sqrt(FP_Rate_final * (1 - FP_Rate_final) / count_Below_AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "No Correction",
  FN_Rate = FN_Rate_final,
  FP_Rate = FP_Rate_final,
  FN_SD = SD_FN,
  FP_SD = SD_FP,
  precision = precision,
  recall = recall,
  F1 = F1,
  f_betal = f_beta
))

```

```{r}
summary_table
```

```{r}
hist(base_conc_sim, 
     main = "Histogram of Base concentration around the AET", 
     xlab = "Concentration",
     col = "skyblue")
hist(simulated_RRF,
     main = "Histogram of RRFs (non-parametric bootstrap of GC/MS dataset 5µg/mL) from FDA Clap list", 
     xlab = "Concentration",
     col = "grey")
```

## GC-MS Correction UF Factor



```{r}
rsd <- function(x) {
  return(sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE))
}


set.seed(123)
N_sim <- 10000
UF_RSD <- 1/(1 - rsd(gcms_clap_clean$"5 µg/mL")) # formula from ISO 10993-18
UF_list <- c(
  "UF (2)"   = 2, 
  "UF (4)"   = 4, 
  "UF (5)"   = 5, 
  "UF (10)"  = 10, 
  "UF (RSD)" = UF_RSD
)

N_extractables <- 80

for (m_name in names(UF_list)) {
  uf_val <- UF_list[[m_name]]
  total_FN <- 0
  total_FP <- 0
  total_TP <- 0
  total_TN <- 0
  
  
for (i in 1:N_sim) {
## 2) Concentration from the log-normal distribution
base_conc_sim <- rlnorm(N_extractables, meanlog = -1.15, sdlog = 1.91)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_extractables, replace = FALSE)
## II) UF corrections
  
  exp_conc <- (base_conc_sim * simulated_RRF)
  corrected_conc <- exp_conc * uf_val 
  
  is_above_threshold <- base_conc_sim > AET
  reported_above_threshold <- corrected_conc > AET

  total_FN <- total_FN + sum(is_above_threshold & !reported_above_threshold, na.rm = TRUE)
  total_FP <- total_FP + sum(!is_above_threshold & reported_above_threshold, na.rm = TRUE)
  total_TP <- total_TP + sum(is_above_threshold & reported_above_threshold, na.rm = TRUE)
  total_TN <- total_TN + sum(!is_above_threshold & !reported_above_threshold, na.rm = TRUE)
}

count_Above_AET <- total_TP + total_FN
count_Below_AET  <- total_TN + total_FP
FN_Rate_final <- total_FN / count_Above_AET
FP_Rate_final <- total_FP / count_Below_AET

recall <- total_TP / count_Above_AET
precision <- total_TP / (total_TP + total_FP)

F1 <- (2 * precision * recall / (precision + recall))
beta <- 5
f_beta <- (1 + beta^2) * (precision * recall) / ((beta^2 * precision) + recall)

SD_FN <- sqrt(FN_Rate_final * (1 - FN_Rate_final) / count_Above_AET)
SD_FP <- sqrt(FP_Rate_final * (1 - FP_Rate_final) / count_Below_AET)



summary_table <- rbind(summary_table, data.frame(
    Method = m_name,
    FN_Rate = FN_Rate_final,
    FP_Rate = FP_Rate_final,
    FN_SD = SD_FN,
    FP_SD = SD_FP,
    precision = precision,
    recall = recall,
    F1 = F1,
    f_betal = f_beta
))
}


UF_RSD
```

```{r}
summary_table
```

## GC-MS Correction Factor percentile

```{r}
percentile <- function(x) {
  return(quantile(x, .16))
}

set.seed(123)
N_sim <- 10000
N_extractables <- 80
total_FN <- 0
total_FP <- 0
total_TP <- 0
total_TN <- 0


for (i in 1:N_sim) {
## 2) Concentration from the log-normal distribution
base_conc_sim <- rlnorm(N_extractables, meanlog = -1.15, sdlog = 1.91)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_extractables, replace = FALSE)

exp_conc <- (base_conc_sim * simulated_RRF)
## III) 16th percentile Correction
corrected_conc <- exp_conc * (mean(simulated_RRF) / percentile(simulated_RRF))

is_above_threshold <- base_conc_sim > AET
reported_above_threshold <- corrected_conc > AET

  total_FN <- total_FN + sum(is_above_threshold & !reported_above_threshold, na.rm = TRUE)
  total_FP <- total_FP + sum(!is_above_threshold & reported_above_threshold, na.rm = TRUE)
  total_TP <- total_TP + sum(is_above_threshold & reported_above_threshold, na.rm = TRUE)
  total_TN <- total_TN + sum(!is_above_threshold & !reported_above_threshold, na.rm = TRUE)
}

count_Above_AET <- total_TP + total_FN
count_Below_AET  <- total_TN + total_FP
FN_Rate_final <- total_FN / count_Above_AET
FP_Rate_final <- total_FP / count_Below_AET

recall <- total_TP / count_Above_AET
precision <- total_TP / (total_TP + total_FP)

F1 <- (2 * precision * recall / (precision + recall))
beta <- 5
f_beta <- (1 + beta^2) * (precision * recall) / ((beta^2 * precision) + recall)

SD_FN <- sqrt(FN_Rate_final * (1 - FN_Rate_final) / count_Above_AET)
SD_FP <- sqrt(FP_Rate_final * (1 - FP_Rate_final) / count_Below_AET)


summary_table <- rbind(summary_table, data.frame(
  Method = "Percentile (16th)",
  FN_Rate = FN_Rate_final,
  FP_Rate = FP_Rate_final,
  FN_SD = SD_FN,
  FP_SD = SD_FP,
  precision = precision,
  recall = recall,
  F1 = F1,
  f_betal = f_beta
))

```

## GC-MS Correction RRFlow

This simulation assumes that we know the true RRF of all compounds sampled and therefore performing the conditional correction. 
We also force the UF to be positive by removing low and high responders. 

```{r}

## RSD is calculated only for RRFs between 0.05 and 1 
## 'Compounds with RRF lower than 0.05 are excluded from UF calculation (20 times lower than the internal standard).page 8
## Therefore, as a criterion for RRF selection, only compounds with an RRF less than 1 are actually considered for UF calculation. page 8

rsd <- function(x) {
  x_use = x[x<=1&x>0.05]
  return(sd(x_use, na.rm = TRUE) / mean(x_use, na.rm = TRUE))
}

## Simulation starts
set.seed(123)
N_sim <- 10000
N_extractables <- 80
total_FN <- 0
total_FP <- 0
total_TP <- 0
total_TN <- 0

# remove 0 values
RRF_GCMS_superior0 <- gcms_clap_clean$"5 µg/mL"
RRF_GCMS_superior0 <- RRF_GCMS_superior0[RRF_GCMS_superior0 > 0]

# Apply Part 18 formula for database
UF <- 1/(1-rsd(RRF_GCMS_superior0))


# Start simulation
for (i in 1:N_sim) {

## 2) Concentration from the log-normal distribution
base_conc_sim <- rlnorm(N_extractables, meanlog = -1.15, sdlog = 1.91)
## A) Sampling of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_extractables, replace = FALSE)
## IV) RRflow condition

exp_conc <- base_conc_sim * simulated_RRF   # experimental concentration : True concentration with the actual RRF
corrected_conc <- exp_conc * UF             # corrected concentration: Exp * UF (conservative method)
condition <- (simulated_RRF < 0.5) | (simulated_RRF > 2) # RRFlow conditions rescaling factor should be focused only to those compounds which exhibits RRF< 0.5 and RRF>2 
corrected_conc[condition] <- exp_conc[condition] / simulated_RRF[condition] # average value of the RRF RRFlow application in extractables amount rescaling ( The average RRF determined by the RRFlow will be used to re-calculate the amount of the corresponding compound detected in the extractables study , will assume that it is in the linearity range) 

  is_above_threshold <- base_conc_sim > AET
  reported_above_threshold <- corrected_conc > AET

  total_FN <- total_FN + sum(is_above_threshold & !reported_above_threshold, na.rm = TRUE)
  total_FP <- total_FP + sum(!is_above_threshold & reported_above_threshold, na.rm = TRUE)
  total_TP <- total_TP + sum(is_above_threshold & reported_above_threshold, na.rm = TRUE)
  total_TN <- total_TN + sum(!is_above_threshold & !reported_above_threshold, na.rm = TRUE)
}

count_Above_AET <- total_TP + total_FN
count_Below_AET  <- total_TN + total_FP
FN_Rate_final <- total_FN / count_Above_AET
FP_Rate_final <- total_FP / count_Below_AET

recall <- total_TP / count_Above_AET
precision <- total_TP / (total_TP + total_FP)

F1 <- (2 * precision * recall / (precision + recall))
beta <- 5
f_beta <- (1 + beta^2) * (precision * recall) / ((beta^2 * precision) + recall)

SD_FN <- sqrt(FN_Rate_final * (1 - FN_Rate_final) / count_Above_AET)
SD_FP <- sqrt(FP_Rate_final * (1 - FP_Rate_final) / count_Below_AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "RRFlow",
  FN_Rate = FN_Rate_final,
  FP_Rate = FP_Rate_final,
  FN_SD = SD_FN,
  FP_SD = SD_FP,
  precision = precision,
  recall = recall,
  F1 = F1,
  f_betal = f_beta
))

```

```{r}
UF
summary_table
```

## GC-MS Correction percentile 10th

Here we divide by the 10th percentile (which is arbitrary) assuming to capture the left hand side of the RRF database (low responders)

```{r}
set.seed(123)
N_sim <- 10000
N_extractables <- 80
total_FN <- 0
total_FP <- 0
total_TP <- 0
total_TN <- 0


RRF_GCMS_superior0 <- gcms_clap_clean$"5 µg/mL"
RRF_GCMS_superior0 <- RRF_GCMS_superior0[RRF_GCMS_superior0 > 0]

conservative_RRF_estimate <- quantile(gcms_clap_clean$"5 µg/mL", 0.10)

for (i in 1:N_sim) {

## 2) Concentration from the log-normal distribution
base_conc_sim <- rlnorm(N_extractables, meanlog = -1.15, sdlog = 1.91)
## A) Non-Parametric Bootstrap of the GC/MS dataset (5µg/mL)
simulated_RRF <- sample(gcms_clap_clean$"5 µg/mL", size = N_extractables, replace = FALSE)
## V RRF < 1

exp_conc <- base_conc_sim * simulated_RRF   # experimental concentration : True concentration with the actual RRF
corrected_conc <- exp_conc / conservative_RRF_estimate 

is_above_threshold <- base_conc_sim > AET
reported_above_threshold <- corrected_conc > AET

  total_FN <- total_FN + sum(is_above_threshold & !reported_above_threshold, na.rm = TRUE)
  total_FP <- total_FP + sum(!is_above_threshold & reported_above_threshold, na.rm = TRUE)
  total_TP <- total_TP + sum(is_above_threshold & reported_above_threshold, na.rm = TRUE)
  total_TN <- total_TN + sum(!is_above_threshold & !reported_above_threshold, na.rm = TRUE)
}

count_Above_AET <- total_TP + total_FN
count_Below_AET  <- total_TN + total_FP
FN_Rate_final <- total_FN / count_Above_AET
FP_Rate_final <- total_FP / count_Below_AET

recall <- total_TP / count_Above_AET
precision <- total_TP / (total_TP + total_FP)


F1 <- (2 * precision * recall / (precision + recall))
beta <- 5
f_beta <- (1 + beta^2) * (precision * recall) / ((beta^2 * precision) + recall)

SD_FN <- sqrt(FN_Rate_final * (1 - FN_Rate_final) / count_Above_AET)
SD_FP <- sqrt(FP_Rate_final * (1 - FP_Rate_final) / count_Below_AET)

summary_table <- rbind(summary_table, data.frame(
  Method = "Percentile 10th",
  FN_Rate = FN_Rate_final,
  FP_Rate = FP_Rate_final,
  FN_SD = SD_FN,
  FP_SD = SD_FP,
  precision = precision,
  recall = recall,
  F1 = F1,
  f_betal = f_beta
))


```

```{r}
summary_table
```

## Optimised percentile


Here we find the best percentile to divide to target a recall of 95% (which is also arbitrary)
```{r}
set.seed(123)

target_recall <- 0.95
found <- FALSE
optimal_percentile <- NA

N_sim <- 10000
N_extractables <- 80

RRF_GCMS_superior0 <- gcms_clap_clean$"5 µg/mL"
RRF_GCMS_superior0 <- RRF_GCMS_superior0[RRF_GCMS_superior0 > 0]

for (test_percentile in seq(0.50, 0.01, by = -0.01)) {
  
  conservative_RRF_estimate <- quantile(RRF_GCMS_superior0, test_percentile, na.rm=TRUE)
  
  total_FN <- 0
  total_FP <- 0
  total_TP <- 0
  total_TN <- 0

  for (i in 1:N_sim) {
    base_conc_sim <- rlnorm(N_extractables, meanlog = -1.15, sdlog = 1.91)
    simulated_RRF <- sample(RRF_GCMS_superior0, size = N_extractables, replace = FALSE)
    
    exp_conc <- base_conc_sim * simulated_RRF   

    corrected_conc <- exp_conc / conservative_RRF_estimate 

    is_above_threshold <- base_conc_sim > AET
    reported_above_threshold <- corrected_conc > AET

    total_FN <- total_FN + sum(is_above_threshold & !reported_above_threshold, na.rm = TRUE)
    total_FP <- total_FP + sum(!is_above_threshold & reported_above_threshold, na.rm = TRUE)
    total_TP <- total_TP + sum(is_above_threshold & reported_above_threshold, na.rm = TRUE)
    total_TN <- total_TN + sum(!is_above_threshold & !reported_above_threshold, na.rm = TRUE)
}


  current_recall <- total_TP / (total_TP + total_FN)
  
  print(paste("Percentile:", test_percentile, "-> Recall:", round(current_recall, 4))) 

  if(current_recall >= target_recall) {
    optimal_percentile <- test_percentile
    found <- TRUE
    break  
  }
}

if(found) {
  print(paste("Optimal Percentile for 95% Recall:", optimal_percentile))
  
  count_Above_AET <- total_TP + total_FN
  count_Below_AET  <- total_TN + total_FP
  
  FN_Rate_final <- total_FN / count_Above_AET
  FP_Rate_final <- total_FP / count_Below_AET

  recall <- total_TP / count_Above_AET
  precision <- total_TP / (total_TP + total_FP)
  
  F1 <- (2 * precision * recall / (precision + recall))
  beta <- 5
  f_beta <- (1 + beta^2) * (precision * recall) / ((beta^2 * precision) + recall)

  SD_FN <- sqrt(FN_Rate_final * (1 - FN_Rate_final) / count_Above_AET)
SD_FP <- sqrt(FP_Rate_final * (1 - FP_Rate_final) / count_Below_AET)

  summary_table <- rbind(summary_table, data.frame(
    Method = paste0("Optimized_P"),
    FN_Rate = FN_Rate_final,
    FP_Rate = FP_Rate_final,
    FN_SD = SD_FN,
    FP_SD = SD_FP,
    precision = precision,
    recall = recall,
    F1 = F1,
    f_betal = f_beta 
  ))
  
} else {
  print("1st percentile was not conservative enough to reach 95% recall.")
}
```


```{r}
summary_table
```




## Machine Learning

Loading the ML Model

```{python}
import joblib
import pandas as pd

# Load the model and the feature list
model = joblib.load('rf_rrf_model.joblib')
feats = joblib.load('features.joblib')

def predict_rrf_raw(df):
    return model.predict(df[feats])
```

```{r}
gc_ms <- read_csv("final_model_data.csv") %>%
  drop_na(BP, `Refractive Index`, TPSA, NHOHCount, NumHDonors, NumRotatableBonds, NOCount) 
```

```{r}
set.seed(123)
N_sim <- 10000
N_extractables <- 80
total_FN <- 0
total_FP <- 0
total_TP <- 0
total_TN <- 0

total_points <- N_sim * N_extractables

for (i in 1:N_sim) {
sim_data <- gc_ms %>% 
  sample_n(N_extractables, replace = TRUE)
  
## 2) Concentration from the log-normal distribution
base_conc_sim <- rlnorm(N_extractables, meanlog = -1.15, sdlog = 1.91)
## VI) RRF generated via the machine learning
simulated_RRF <- py$predict_rrf_raw(sim_data)
## no correction
corrected_conc <- base_conc_sim * simulated_RRF

is_above_threshold <- base_conc_sim > AET
reported_above_threshold <- corrected_conc > AET

  total_FN <- total_FN + sum(is_above_threshold & !reported_above_threshold, na.rm = TRUE)
  total_FP <- total_FP + sum(!is_above_threshold & reported_above_threshold, na.rm = TRUE)
  total_TP <- total_TP + sum(is_above_threshold & reported_above_threshold, na.rm = TRUE)
  total_TN <- total_TN + sum(!is_above_threshold & !reported_above_threshold, na.rm = TRUE)
}

count_Above_AET <- total_TP + total_FN
count_Below_AET  <- total_TN + total_FP
FN_Rate_final <- total_FN / count_Above_AET
FP_Rate_final <- total_FP / count_Below_AET

recall <- total_TP / count_Above_AET
precision <- total_TP / (total_TP + total_FP)

F1 <- (2 * precision * recall / (precision + recall))
beta <- 5
f_beta <- (1 + beta^2) * (precision * recall) / ((beta^2 * precision) + recall)

SD_FN <- sqrt(FN_Rate_final * (1 - FN_Rate_final) / count_Above_AET)
SD_FP <- sqrt(FP_Rate_final * (1 - FP_Rate_final) / count_Below_AET)


summary_table <- rbind(summary_table, data.frame(
  Method = "Machine Learning",
  FN_Rate = FN_Rate_final,
  FP_Rate = FP_Rate_final,
  FN_SD = SD_FN,
  FP_SD = SD_FP,
  precision = precision,
  recall = recall,
  F1 = F1,
  f_betal = f_beta
))

```

```{r}
summary_table
```



## Comparison

```{r}
plot_data <- summary_table %>%
  pivot_longer(cols = c(FN_Rate, FP_Rate), names_to = "Type", values_to = "Rate") %>%
  mutate(
    SD = ifelse(Type == "FN_Rate", round(FN_SD*100, 2), round(FP_SD*100, 2)),
    Type = ifelse(Type == "FN_Rate", "False Negative", "False Positive")
  )

method_levels <- c(
  "No Correction",
  "UF (2)", 
  "UF (4)", 
  "UF (5)", 
  "UF (10)",  
  "UF (RSD)",
  "Percentile (16th)", 
  "RRFlow",
  "Percentile 10th",
  "Machine Learning",
  "Optimized_P")

plot_data$Method <- factor(plot_data$Method, levels = method_levels)


ggplot(plot_data, aes(x = Method, y = Rate*100, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), alpha = 0.8) +
  geom_errorbar(aes(ymin = pmax(0, Rate*100 - 1.96*SD), ymax = Rate*100 + 1.96*SD), 
                position = position_dodge(width = 0.9), width = 0.2) +
  labs(
    title = paste0("Comparison of GC-MS Correction Methods (5 µg/mL)", " AET=", AET, "µg/mL"),
    subtitle = "Conc. from Log/Normal Distribution, RRF sampled from CLAP list data",
    y = "Rate (%)",
    x = "Method") +
  geom_text(aes(label = sprintf("%.1f%%", Rate*100)), position = position_dodge(width = 0.9), hjust = -0.5, size = 3, vjust = 0.1) +
  scale_fill_manual(values = c("False Negative" = "red", "False Positive" = "grey")) + 
  scale_y_continuous(limits = c(0, 100)) +
  theme_minimal() +
  coord_flip()

```

```{r}
summary_table %>%
  select(Method, precision, recall, F1, f_betal)
```

ROC and Precision Recall plots

```{r}
ggplot(summary_table, aes(x = FP_Rate, y = recall, color = Method)) +
  geom_point(size = 4) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") + 
  xlim(0, 1) + ylim(0, 1) +
  labs(title = "ROC",
       x = "False Positive Rate (1-Specificity)",
       y = "Recall (Sensitivity)") +
  theme_minimal()

ggplot(summary_table, aes(x = recall, y = precision, color = Method)) +
  geom_point(size = 4) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") + # Random guess line
  xlim(0, 1) + ylim(0, 1) +
  labs(title = "Precision/Recall",
       x = "Recall",
       y = "Precision") +
  theme_minimal()

```


