---
title: "el_bias"
format: gfm
editor: visual
---

# Libraries

```{r}

# Load required packages
library(tidyverse)
library(lubridate)
library(ggrepel)
library(here)
library(parallel)
library(readxl)
library(reticulate)
library(MASS)

```

```{r}
#| include: false
# Set global plot settings for PNG output on Linux
knitr::opts_chunk$set(
  dev = "png", 
  dpi = 300, 
  dev.args = list(type = "cairo")
) 
```

# Data Extraction

Data for RRFs are coming from the US FDA Clap List. <https://cdrh-rst.fda.gov/chemicals-list-analytical-performance-clap>

Each file is in a pdf format and contains RRF values for 3 concentration of a surrogate. LCMS RRFs CLAP Negative Ion Detection.pdf (255.54 KB) LCMS RRFs CLAP Positive Ion Detection.pdf (279.48 KB) RRF GCMS for CLAP.pdf (322.77 KB)

They are first extracted to perform the bootstrap simulation. Some data are n.d. or below DL. The current strategy is to change the n.d. to 0 and below DL to 0.001 based on the input of the Subject Matter Expert.

# GC/MS data

```{r}
gcms_clap <- read_excel( 
  path = "processed_data/GCMS.xlsx", 
  sheet = "GC-MS")

gcms_clap_clean <- gcms_clap %>%
  na.omit()

gcms_clap_clean

gcms_clap_clean$"5 µg/mL" <- as.numeric(gcms_clap_clean$"5 µg/mL")
gcms_clap_clean$"10 µg/mL" <- as.numeric(gcms_clap_clean$"10 µg/mL")
gcms_clap_clean$"20 µg/mL" <- as.numeric(gcms_clap_clean$"20 µg/mL")

hist(gcms_clap_clean$"5 µg/mL", 
     main = "Histogram of RRFs for LC/MS with a 5 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "skyblue")
hist(gcms_clap_clean$"10 µg/mL", 
     main = "Histogram of RRFs for LC/MS with a 10 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "orange")
hist(gcms_clap_clean$"20 µg/mL", 
     main = "Histogram of RRFs for LC/MS with a 20 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "red")

```

# LC/MS pos data

```{r}
lcmspos_clap <- read_excel( 
  path = "processed_data/LCMS.xlsx", 
  sheet = "LCMSPOS")

lcmspos_clap_clean <- lcmspos_clap %>%
  na.omit()


hist(lcmspos_clap_clean$"5 µg/mL", 
     main = "Histogram of RRFs for LC/MS + with a 5 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "skyblue")
hist(lcmspos_clap_clean$"10 µg/mL", 
     main = "Histogram of RRFs for LC/MS + with a 10 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "orange")
hist(lcmspos_clap_clean$"20 µg/mL", 
     main = "Histogram of RRFs for LC/MS + with a 20 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "red")

```

# LC/MS neg data

```{r}
lcmsneg_clap <- read_excel( 
  path = "processed_data/LCMS.xlsx", 
  sheet = "LCMSNEG")

lcmsneg_clap_clean <- lcmsneg_clap %>%
  na.omit()

lcmsneg_clap_clean


hist(lcmsneg_clap_clean$"5 µg/mL", 
     main = "Histogram of RRFs for LC/MS- with a 5 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "skyblue")
hist(lcmsneg_clap_clean$"10 µg/mL", 
     main = "Histogram of RRFs for LC/MS- with a 10 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "orange")
hist(lcmsneg_clap_clean$"20 µg/mL", 
     main = "Histogram of RRFs for LC/MS- with a 20 µg/mL surrogate Concentration", 
     xlab = "RRF",
     col = "red")

```

# Generate Base concentration

## Define AET

```{r}
AET <- 1
```

## Uniform distribution

```{r}
set.seed(123)
base_concentrations1 <- runif(100, min = 0.5 * AET, max = 2 * AET)
hist(base_concentrations1)
```

## Log-Normal distribution

$log(0.001) = -6.907, log(100) = 4.605$ \\ $mid (meanlog) approx -1.15$ \\ $width = 4.605 - (-6.907) = 11.51$ \\ $width/6 = 11.51/6=1.91$

```{r}
set.seed(123)
base_concentrations2 <- rlnorm(100, meanlog = -1.15, sdlog = 1.91)
hist(base_concentrations2)
```

## Custom distribution

```{r}
set.seed(123)
Z <- runif(70, min = 0.001, max = 1)
Y <- runif(20, min = 1, max = 10)
K <- runif(10, min = 10, max = 100)

custom <- c(Z, Y, K)

custom_sampled <- sample(custom, size = 100, replace = FALSE)

hist(custom_sampled)
```

## Normal distribution (simulating a GC/MS RRF distribution)

```{r}
set.seed(123)
RRF_norm <- rnorm(50, 1, 0.2)
RRF_norm[RRF_norm<0] <- 0.001 # avoiding negative numbers
hist(RRF_norm)
```

## Log_Normal distribution (simulating a LC/MS RRF distribution)

```{r}
set.seed(123)
RRF_lnorm <- rlnorm(50, 0, 0.5)
RRF_lnorm[RRF_lnorm<0] <- 0.001 # avoiding negative numbers
hist(RRF_lnorm)
```

# Fit the distribution with fitdistr

```{r}
set.seed(123)

RRF_superior0 <- lcmsneg_clap_clean$"5 µg/mL"
RRF_superior0 <- RRF_superior0[RRF_superior0 > 0]
fit_normal <- fitdistr(RRF_superior0, "normal")
print(fit_normal)
fit_lognormal <- fitdistr(RRF_superior0, "lognormal")
print(fit_lognormal)
fit_gamma <- fitdistr(RRF_superior0, "gamma")
print(fit_gamma)

```

```{r}
RRF_df <- data.frame(RRF = RRF_superior0)

norm_mean <- fit_normal$estimate["mean"]
norm_sd   <- fit_normal$estimate["sd"]
lnorm_meanlog <- fit_lognormal$estimate["meanlog"]
lnorm_sdlog   <- fit_lognormal$estimate["sdlog"]
gamma_shape <- fit_gamma$estimate["shape"]
gamma_rate  <- fit_gamma$estimate["rate"]


ggplot(RRF_df, aes(x = RRF)) +
  geom_histogram(aes(y = after_stat(density)), 
                 bins = 20, 
                 fill = "lightgray", 
                 color = "white") +
  stat_function(fun = dnorm, 
                args = list(mean = norm_mean, sd = norm_sd),
                aes(color = "Normal", linetype = "Normal"), 
                linewidth = 1) +
  stat_function(fun = dlnorm, 
                args = list(meanlog = lnorm_meanlog, sdlog = lnorm_sdlog),
                aes(color = "Log-normal", linetype = "Log-normal"), 
                linewidth = 1) +
  stat_function(fun = dgamma, 
                args = list(shape = gamma_shape, rate = gamma_rate),
                aes(color = "Gamma", linetype = "Gamma"), 
                linewidth = 1) +
  scale_color_manual(name = "Distribution",
                     values = c("Normal" = "darkblue", 
                                "Log-normal" = "darkred", 
                                "Gamma" = "darkgreen")) +
  scale_linetype_manual(name = "Distribution",
                        values = c("Normal" = "solid", 
                                   "Log-normal" = "dashed", 
                                   "Gamma" = "dotted")) +
  labs(title = "Fit Comparison: Normal, Log-normal, and Gamma",
       x = "RRF",
       y = "Density") +
  theme_minimal()

```

```{r}
aic_normal <- AIC(fit_normal)
aic_lognormal <- AIC(fit_lognormal)
aic_gamma <- AIC(fit_gamma)

aic_normal
aic_lognormal
aic_gamma
```

```{r}
library(patchwork)

# 1. Normal Q-Q Plot
qq_norm <- ggplot(RRF_df, aes(sample = RRF)) +
  stat_qq(distribution = qnorm,
          dparams = list(mean = norm_mean, sd = norm_sd),
          alpha = 0.5) +
  stat_qq_line(distribution = qnorm,
               dparams = list(mean = norm_mean, sd = norm_sd),
               color = "darkblue", linewidth = 1) +
  labs(title = "Normal Q-Q", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()

# 2. Log-normal Q-Q Plot
qq_lnorm <- ggplot(RRF_df, aes(sample = RRF)) +
  stat_qq(distribution = qlnorm,
          dparams = list(meanlog = lnorm_meanlog, sdlog = lnorm_sdlog),
          alpha = 0.5) +
  stat_qq_line(distribution = qlnorm,
               dparams = list(meanlog = lnorm_meanlog, sdlog = lnorm_sdlog),
               color = "darkred", linewidth = 1) +
  labs(title = "Log-normal Q-Q", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()

# 3. Gamma Q-Q Plot
qq_gamma <- ggplot(RRF_df, aes(sample = RRF)) +
  stat_qq(distribution = qgamma,
          dparams = list(shape = gamma_shape, rate = gamma_rate),
          alpha = 0.5) +
  stat_qq_line(distribution = qgamma,
               dparams = list(shape = gamma_shape, rate = gamma_rate),
               color = "darkgreen", linewidth = 1) +
  labs(title = "Gamma Q-Q", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()

# Display them side-by-side
qq_norm + qq_lnorm + qq_gamma
```

# Scenario with uniform distribution & GC/MS data from US FDA CLAP list

## Initialisation of the comparison table/

```{r}
summary_table <- data.frame(
  Method = character(),
  FN_Rate = numeric(),
  FP_Rate = numeric(),
  FN_SD = numeric(),
  FP_SD = numeric(),
  precision = numeric(),
  recall = numeric(),
  F1 = numeric(),
  f_beta = numeric(),
  Youden_Y = numeric(),
  Bias_Q = numeric(),
  Efficiency_EI = numeric()
)

all_distributions <- data.frame()

```

## LC-MS No Correction Scenarios

```{r}
set.seed(123)
N_sim <- 10000
dist_Y <- numeric(N_sim)

dist_csi <- numeric(N_sim)
dist_psi <- numeric(N_sim)
dist_sui <- numeric(N_sim)
dist_pcui <- numeric(N_sim)
dist_ncui <- numeric(N_sim)
dist_ActualPos <- numeric(N_sim)
dist_ActualNeg <- numeric(N_sim)

dist_EI <- numeric(N_sim)
dist_Q <- numeric(N_sim)
dist_MCC <- numeric(N_sim)
dist_DOR <- numeric(N_sim)
N_extractables <- 80
total_FN <- 0
total_FP <- 0
total_TP <- 0
total_TN <- 0
dist_f1 <- numeric(N_sim)
dist_f5 <- numeric(N_sim)
dist_fnr <- numeric(N_sim)
dist_fpr <- numeric(N_sim)
beta <- 5

for (i in 1:N_sim) {
## 1) Concentration around the AET
base_conc_sim <- runif(N_extractables, min = 0.5 * AET, max = 2 * AET)
## A) Sampling from the LC/MS- dataset fitted to gamma dist (5µg/mL – fitted distribution from CLAP list) (5µg/mL)
simulated_RRF <- rgamma(N_extractables, gamma_shape, gamma_rate)
## I) No Correction
corrected_conc <- base_conc_sim * simulated_RRF

is_at_and_above_threshold <- base_conc_sim >= AET
reported_at_and_above_threshold <- corrected_conc >= AET

## Count per iteration
  iter_TP <- sum(is_at_and_above_threshold & reported_at_and_above_threshold, na.rm = TRUE)
  iter_FP <- sum(!is_at_and_above_threshold & reported_at_and_above_threshold, na.rm = TRUE)
  iter_FN <- sum(is_at_and_above_threshold & !reported_at_and_above_threshold, na.rm = TRUE)
  iter_TN <- sum(!is_at_and_above_threshold & !reported_at_and_above_threshold, na.rm = TRUE)

  total_TP <- total_TP + iter_TP
  total_FP <- total_FP + iter_FP
  total_FN <- total_FN + iter_FN
  total_TN <- total_TN + iter_TN

  # Matthews Correlation Coefficient (MCC)
  denom_MCC <- sqrt(as.numeric(iter_TP+iter_FP)*(iter_TP+iter_FN)*(iter_TN+iter_FP)*(iter_TN+iter_FN))
  iter_MCC <- if(denom_MCC > 0) ((as.numeric(iter_TP)*iter_TN) - (as.numeric(iter_FP)*iter_FN)) / denom_MCC else 0
  
  # Diagnostic Odds Ratio (DOR)
  iter_DOR <- (as.numeric(iter_TP) * iter_TN) / ((as.numeric(iter_FP) * iter_FN) + 1e-6)
  
  dist_MCC[i] <- iter_MCC
  dist_DOR[i] <- iter_DOR


  # metrics per iterations
  iter_prec <- if((iter_TP + iter_FP) > 0) iter_TP / (iter_TP + iter_FP) else 0
  iter_rec  <- if((iter_TP + iter_FN) > 0) iter_TP / (iter_TP + iter_FN) else 0
  
  iter_ActualPos <- iter_TP + iter_FN
  iter_ActualNeg <- iter_TN + iter_FP

  iter_fnr <- if(iter_ActualPos > 0) iter_FN / iter_ActualPos else 0
  iter_fpr <- if(iter_ActualNeg > 0) iter_FP / iter_ActualNeg else 0
  
  dist_fnr[i] <- iter_fnr * 100
  dist_fpr[i] <- iter_fpr * 100
  
  # F1
  iter_f1 <- if((iter_prec + iter_rec) > 0) {
    (2 * iter_prec * iter_rec) / (iter_prec + iter_rec)
  } else { 0 }
  
  # F5
  denom_beta <- (beta^2 * iter_prec) + iter_rec
  iter_fbeta <- if(denom_beta > 0) {
    ((1 + beta^2) * (iter_prec * iter_rec)) / denom_beta
  } else { 0 }
  
  dist_f1[i]   <- iter_f1
  dist_f5[i]   <- iter_fbeta
  iter_ActualNeg <- iter_TN + iter_FP
  iter_Total     <- iter_TP + iter_FP + iter_FN + iter_TN
  iter_spec      <- if(iter_ActualNeg > 0) iter_TN / iter_ActualNeg else 0

  # Larner's Metrics
  iter_Y    <- iter_rec + iter_spec - 1 # Youden Index
  iter_UAcc <- (iter_rec + iter_spec) / 2 # Unbiased Accuracy
  iter_Q    <- (iter_TP + iter_FP) / iter_Total # Bias / Positive Sign Rate
  iter_EI   <- (iter_TP + iter_TN - (iter_FP + iter_FN)) / iter_Total # Efficiency Index

  dist_Y[i]  <- iter_Y
  dist_EI[i] <- iter_EI
  dist_Q[i]  <- iter_Q

  # --- Calculation of Larner's Metrics (add inside the simulation loop) ---

  # Pre-calculate base probabilities for indices
  iter_Sens <- if(iter_ActualPos > 0) iter_TP / iter_ActualPos else 0
  iter_Spec <- if(iter_ActualNeg > 0) iter_TN / iter_ActualNeg else 0
  iter_PPV  <- if((iter_TP + iter_FP) > 0) iter_TP / (iter_TP + iter_FP) else 0
  iter_NPV  <- if((iter_TN + iter_FN) > 0) iter_TN / (iter_TN + iter_FN) else 0

  # 1. Critical Success Index (CSI) - Ignores True Negatives
  iter_CSI <- if((iter_TP + iter_FN + iter_FP) > 0) {
    iter_TP / (iter_TP + iter_FN + iter_FP) 
  } else { 0 }

  # 2. Predictive Summary Index (PSI) - Net gain of certainty
  iter_PSI <- iter_PPV + iter_NPV - 1

  # 3. Clinical Utility Indexes (PCUI & NCUI)
  iter_PCUI <- iter_Sens * iter_PPV
  iter_NCUI <- iter_Spec * iter_NPV

  # 4. Summary Utility Index (SUI) - Unitary score of utility 
  iter_SUI <- iter_PCUI + iter_NCUI

  dist_csi[i] <- iter_CSI
  dist_psi[i] <- iter_PSI
  dist_sui[i] <- iter_SUI
  dist_pcui[i] <- iter_PCUI
  dist_ncui[i] <- iter_NCUI
  dist_ActualPos[i] <- iter_ActualPos
  dist_ActualNeg[i] <- iter_ActualNeg
}

count_Above_AET <- total_TP + total_FN
count_Below_AET  <- total_TN + total_FP
FN_Rate_final <- total_FN / count_Above_AET
FP_Rate_final <- total_FP / count_Below_AET

recall <- total_TP / count_Above_AET
precision <- total_TP / (total_TP + total_FP)

F1 <- (2 * precision * recall / (precision + recall))
beta <- 5
f_beta <- (1 + beta^2) * (precision * recall) / ((beta^2 * precision) + recall)

SD_FN <- sqrt(FN_Rate_final * (1 - FN_Rate_final) / count_Above_AET)
SD_FP <- sqrt(FP_Rate_final * (1 - FP_Rate_final) / count_Below_AET)

summary_table <- bind_rows(
  summary_table,
  data.frame(
    Method = "No Correction",
    FN_Rate = FN_Rate_final,
    FP_Rate = FP_Rate_final,
    FN_SD = SD_FN,
    FP_SD = SD_FP,
    precision = precision,
    recall = recall,
    F1 = F1,
    f_beta = f_beta,
    Youden_Y = mean(dist_Y, na.rm = TRUE),
    Bias_Q = mean(dist_Q, na.rm = TRUE),
    Efficiency_EI = mean(dist_EI, na.rm = TRUE),
    MCC = mean(dist_MCC, na.rm = TRUE),
    DOR = mean(dist_DOR, na.rm = TRUE),
    
    # New Larner Metrics
    CSI = mean(dist_csi, na.rm = TRUE),
    PSI = mean(dist_psi, na.rm = TRUE),
    SUI = mean(dist_sui, na.rm = TRUE),
    PCUI = mean(dist_pcui, na.rm = TRUE),
    NCUI = mean(dist_ncui, na.rm = TRUE),

    # Absolute Counts and UF
    is_at_and_above_threshold = count_Above_AET,
    X.is_at_and_above_threshold = count_Below_AET,
    reported_at_and_above_threshold = total_TP + total_FP,
    X.reported_at_and_above_threshold = total_TN + total_FN,
    perc_is_at_and_above_threshold = (count_Above_AET / (count_Above_AET + count_Below_AET)) * 100,
    perc_reported_at_and_above_threshold = ((total_TP + total_FP) / (count_Above_AET + count_Below_AET)) * 100,
    UF = 1
  )
)

metrics_df <- data.frame(
  Method = "No Correction",
  F1 = dist_f1,
  F5 = dist_f5,
  Percent_FN = dist_fnr,
  Percent_FP = dist_fpr,
  EI = dist_EI,
  Youden_Y = dist_Y,
  Bias_Q = dist_Q,
  MCC = dist_MCC,
  DOR = dist_DOR
)

all_distributions <- rbind(all_distributions, metrics_df)

```

```{r}
hist(simulated_RRF)
```

```{r}
hist(base_conc_sim, 
     main = "Histogram of Base concentration around the AET", 
     xlab = "Concentration",
     col = "skyblue")
hist(simulated_RRF,
     main = "Histogram of RRFs (non-parametric bootstrap of GC/MS dataset 5µg/mL) from FDA Clap list", 
     xlab = "Concentration",
     col = "grey")
```

## LC-MS Correction UF Factor

```{r}
rsd <- function(x) {
  return(sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE))
}


set.seed(123)
N_sim <- 10000
dist_Y <- numeric(N_sim)

dist_csi <- numeric(N_sim)
dist_psi <- numeric(N_sim)
dist_sui <- numeric(N_sim)
dist_pcui <- numeric(N_sim)
dist_ncui <- numeric(N_sim)
dist_ActualPos <- numeric(N_sim)
dist_ActualNeg <- numeric(N_sim)

dist_EI <- numeric(N_sim)
dist_Q <- numeric(N_sim)
dist_MCC <- numeric(N_sim)
dist_DOR <- numeric(N_sim)
UF_RSD <- 1/(1-rsd(rgamma(N_extractables, gamma_shape, gamma_rate))) # formula from ISO 10993-18
UF_list <- c(
  "UF (2)"   = 2, 
  "UF (4)"   = 4, 
  "UF (5)"   = 5, 
  "UF (10)"  = 10, 
  "UF (RSD)" = UF_RSD
)

N_extractables <- 80

for (m_name in names(UF_list)) {
  uf_val <- UF_list[[m_name]]
  total_FN <- 0
  total_FP <- 0
  total_TP <- 0
  total_TN <- 0
  
  
for (i in 1:N_sim) {
## 1) Concentration around the AET
base_conc_sim <- runif(N_extractables, min = 0.5 * AET, max = 2 * AET)
## A) Sampling from the LC/MS- dataset fitted to gamma dist (5µg/mL – fitted distribution from CLAP list) (5µg/mL)
simulated_RRF <- rgamma(N_extractables, gamma_shape, gamma_rate)
## II) UF corrections
  
  exp_conc <- (base_conc_sim * simulated_RRF)
  corrected_conc <- exp_conc * uf_val 
  
  is_at_and_above_threshold <- base_conc_sim >= AET
  reported_at_and_above_threshold <- corrected_conc >= AET

## Count per iteration
  iter_TP <- sum(is_at_and_above_threshold & reported_at_and_above_threshold, na.rm = TRUE)
  iter_FP <- sum(!is_at_and_above_threshold & reported_at_and_above_threshold, na.rm = TRUE)
  iter_FN <- sum(is_at_and_above_threshold & !reported_at_and_above_threshold, na.rm = TRUE)
  iter_TN <- sum(!is_at_and_above_threshold & !reported_at_and_above_threshold, na.rm = TRUE)

  total_TP <- total_TP + iter_TP
  total_FP <- total_FP + iter_FP
  total_FN <- total_FN + iter_FN
  total_TN <- total_TN + iter_TN

  # Matthews Correlation Coefficient (MCC)
  denom_MCC <- sqrt(as.numeric(iter_TP+iter_FP)*(iter_TP+iter_FN)*(iter_TN+iter_FP)*(iter_TN+iter_FN))
  iter_MCC <- if(denom_MCC > 0) ((as.numeric(iter_TP)*iter_TN) - (as.numeric(iter_FP)*iter_FN)) / denom_MCC else 0
  
  # Diagnostic Odds Ratio (DOR)
  iter_DOR <- (as.numeric(iter_TP) * iter_TN) / ((as.numeric(iter_FP) * iter_FN) + 1e-6)
  
  dist_MCC[i] <- iter_MCC
  dist_DOR[i] <- iter_DOR


  # metrics per iterations
  iter_prec <- if((iter_TP + iter_FP) > 0) iter_TP / (iter_TP + iter_FP) else 0
  iter_rec  <- if((iter_TP + iter_FN) > 0) iter_TP / (iter_TP + iter_FN) else 0
  
  iter_ActualPos <- iter_TP + iter_FN
  iter_ActualNeg <- iter_TN + iter_FP

  iter_fnr <- if(iter_ActualPos > 0) iter_FN / iter_ActualPos else 0
  iter_fpr <- if(iter_ActualNeg > 0) iter_FP / iter_ActualNeg else 0
  
  dist_fnr[i] <- iter_fnr * 100
  dist_fpr[i] <- iter_fpr * 100
  
  # F1
  iter_f1 <- if((iter_prec + iter_rec) > 0) {
    (2 * iter_prec * iter_rec) / (iter_prec + iter_rec)
  } else { 0 }
  
  # F5
  denom_beta <- (beta^2 * iter_prec) + iter_rec
  iter_fbeta <- if(denom_beta > 0) {
    ((1 + beta^2) * (iter_prec * iter_rec)) / denom_beta
  } else { 0 }
  
  dist_f1[i]   <- iter_f1
  dist_f5[i]   <- iter_fbeta
  iter_ActualNeg <- iter_TN + iter_FP
  iter_Total     <- iter_TP + iter_FP + iter_FN + iter_TN
  iter_spec      <- if(iter_ActualNeg > 0) iter_TN / iter_ActualNeg else 0

  # Larner's Metrics
  iter_Y    <- iter_rec + iter_spec - 1 # Youden Index
  iter_UAcc <- (iter_rec + iter_spec) / 2 # Unbiased Accuracy
  iter_Q    <- (iter_TP + iter_FP) / iter_Total # Bias / Positive Sign Rate
  iter_EI   <- (iter_TP + iter_TN - (iter_FP + iter_FN)) / iter_Total # Efficiency Index

  dist_Y[i]  <- iter_Y
  dist_EI[i] <- iter_EI
  dist_Q[i]  <- iter_Q

  # --- Calculation of Larner's Metrics (add inside the simulation loop) ---

  # Pre-calculate base probabilities for indices
  iter_Sens <- if(iter_ActualPos > 0) iter_TP / iter_ActualPos else 0
  iter_Spec <- if(iter_ActualNeg > 0) iter_TN / iter_ActualNeg else 0
  iter_PPV  <- if((iter_TP + iter_FP) > 0) iter_TP / (iter_TP + iter_FP) else 0
  iter_NPV  <- if((iter_TN + iter_FN) > 0) iter_TN / (iter_TN + iter_FN) else 0

  # 1. Critical Success Index (CSI) - Ignores True Negatives
  iter_CSI <- if((iter_TP + iter_FN + iter_FP) > 0) {
    iter_TP / (iter_TP + iter_FN + iter_FP) 
  } else { 0 }

  # 2. Predictive Summary Index (PSI) - Net gain of certainty
  iter_PSI <- iter_PPV + iter_NPV - 1

  # 3. Clinical Utility Indexes (PCUI & NCUI)
  iter_PCUI <- iter_Sens * iter_PPV
  iter_NCUI <- iter_Spec * iter_NPV

  # 4. Summary Utility Index (SUI) - Unitary score of utility 
  iter_SUI <- iter_PCUI + iter_NCUI

  dist_csi[i] <- iter_CSI
  dist_psi[i] <- iter_PSI
  dist_sui[i] <- iter_SUI
  dist_pcui[i] <- iter_PCUI
  dist_ncui[i] <- iter_NCUI
  dist_ActualPos[i] <- iter_ActualPos
  dist_ActualNeg[i] <- iter_ActualNeg
}

count_Above_AET <- total_TP + total_FN
count_Below_AET  <- total_TN + total_FP
FN_Rate_final <- total_FN / count_Above_AET
FP_Rate_final <- total_FP / count_Below_AET

recall <- total_TP / count_Above_AET
precision <- total_TP / (total_TP + total_FP)

F1 <- (2 * precision * recall / (precision + recall))
beta <- 5
f_beta <- (1 + beta^2) * (precision * recall) / ((beta^2 * precision) + recall)

SD_FN <- sqrt(FN_Rate_final * (1 - FN_Rate_final) / count_Above_AET)
SD_FP <- sqrt(FP_Rate_final * (1 - FP_Rate_final) / count_Below_AET)

summary_table <- bind_rows(
  summary_table,
  data.frame(
    Method = m_name,
    FN_Rate = FN_Rate_final,
    FP_Rate = FP_Rate_final,
    FN_SD = SD_FN,
    FP_SD = SD_FP,
    precision = precision,
    recall = recall,
    F1 = F1,
    f_beta = f_beta,
    Youden_Y = mean(dist_Y, na.rm = TRUE),
    Bias_Q = mean(dist_Q, na.rm = TRUE),
    Efficiency_EI = mean(dist_EI, na.rm = TRUE),
    MCC = mean(dist_MCC, na.rm = TRUE),
    DOR = mean(dist_DOR, na.rm = TRUE),
    
    # New Larner Metrics
    CSI = mean(dist_csi, na.rm = TRUE),
    PSI = mean(dist_psi, na.rm = TRUE),
    SUI = mean(dist_sui, na.rm = TRUE),
    PCUI = mean(dist_pcui, na.rm = TRUE),
    NCUI = mean(dist_ncui, na.rm = TRUE),

    # Absolute Counts and UF
    is_at_and_above_threshold = count_Above_AET,
    X.is_at_and_above_threshold = count_Below_AET,
    reported_at_and_above_threshold = total_TP + total_FP,
    X.reported_at_and_above_threshold = total_TN + total_FN,
    perc_is_at_and_above_threshold = (count_Above_AET / (count_Above_AET + count_Below_AET)) * 100,
    perc_reported_at_and_above_threshold = ((total_TP + total_FP) / (count_Above_AET + count_Below_AET)) * 100,
    UF = uf_val
  )
)

metrics_df <- data.frame(
  Method = m_name,
  F1 = dist_f1,
  F5 = dist_f5,
  Percent_FN = dist_fnr,
  Percent_FP = dist_fpr,
  EI = dist_EI,
  Youden_Y = dist_Y,
  Bias_Q = dist_Q,
  MCC = dist_MCC,
  DOR = dist_DOR
)

all_distributions <- rbind(all_distributions, metrics_df)

}

UF_RSD

```

```{r}
summary_table
```

## LC-MS Correction Factor mean/16th percentile

```{r}
percentile <- function(x) {
  return(quantile(x, .16))
}

set.seed(123)
N_sim <- 10000
dist_Y <- numeric(N_sim)

dist_csi <- numeric(N_sim)
dist_psi <- numeric(N_sim)
dist_sui <- numeric(N_sim)
dist_pcui <- numeric(N_sim)
dist_ncui <- numeric(N_sim)
dist_ActualPos <- numeric(N_sim)
dist_ActualNeg <- numeric(N_sim)

dist_EI <- numeric(N_sim)
dist_Q <- numeric(N_sim)
dist_MCC <- numeric(N_sim)
dist_DOR <- numeric(N_sim)
N_extractables <- 80
total_FN <- 0
total_FP <- 0
total_TP <- 0
total_TN <- 0

RRF_LCMS_superior0 <- lcmsneg_clap_clean$"5 µg/mL"
RRF_LCMS_superior0 <- RRF_LCMS_superior0[RRF_LCMS_superior0 > 0]

meanLCMS <- mean(rgamma(N_extractables, gamma_shape, gamma_rate))
percentileLCMS <- percentile(rgamma(N_extractables, gamma_shape, gamma_rate))

for (i in 1:N_sim) {
## 1) Concentration around the AET
base_conc_sim <- runif(N_extractables, min = 0.5 * AET, max = 2 * AET)
## A) Sampling from the LC/MS- dataset fitted to gamma dist (5µg/mL – fitted distribution from CLAP list) (5µg/mL)
simulated_RRF <- rgamma(N_extractables, gamma_shape, gamma_rate)

exp_conc <- (base_conc_sim * simulated_RRF)
## III) 16th percentile Correction
corrected_conc <- exp_conc * (meanLCMS / percentileLCMS)

is_at_and_above_threshold <- base_conc_sim >= AET
reported_at_and_above_threshold <- corrected_conc >= AET

## Count per iteration
  iter_TP <- sum(is_at_and_above_threshold & reported_at_and_above_threshold, na.rm = TRUE)
  iter_FP <- sum(!is_at_and_above_threshold & reported_at_and_above_threshold, na.rm = TRUE)
  iter_FN <- sum(is_at_and_above_threshold & !reported_at_and_above_threshold, na.rm = TRUE)
  iter_TN <- sum(!is_at_and_above_threshold & !reported_at_and_above_threshold, na.rm = TRUE)

  total_TP <- total_TP + iter_TP
  total_FP <- total_FP + iter_FP
  total_FN <- total_FN + iter_FN
  total_TN <- total_TN + iter_TN

  # Matthews Correlation Coefficient (MCC)
  denom_MCC <- sqrt(as.numeric(iter_TP+iter_FP)*(iter_TP+iter_FN)*(iter_TN+iter_FP)*(iter_TN+iter_FN))
  iter_MCC <- if(denom_MCC > 0) ((as.numeric(iter_TP)*iter_TN) - (as.numeric(iter_FP)*iter_FN)) / denom_MCC else 0
  
  # Diagnostic Odds Ratio (DOR)
  iter_DOR <- (as.numeric(iter_TP) * iter_TN) / ((as.numeric(iter_FP) * iter_FN) + 1e-6)
  
  dist_MCC[i] <- iter_MCC
  dist_DOR[i] <- iter_DOR


  # metrics per iterations
  iter_prec <- if((iter_TP + iter_FP) > 0) iter_TP / (iter_TP + iter_FP) else 0
  iter_rec  <- if((iter_TP + iter_FN) > 0) iter_TP / (iter_TP + iter_FN) else 0
  
  iter_ActualPos <- iter_TP + iter_FN
  iter_ActualNeg <- iter_TN + iter_FP

  iter_fnr <- if(iter_ActualPos > 0) iter_FN / iter_ActualPos else 0
  iter_fpr <- if(iter_ActualNeg > 0) iter_FP / iter_ActualNeg else 0
  
  dist_fnr[i] <- iter_fnr * 100
  dist_fpr[i] <- iter_fpr * 100
  
  # F1
  iter_f1 <- if((iter_prec + iter_rec) > 0) {
    (2 * iter_prec * iter_rec) / (iter_prec + iter_rec)
  } else { 0 }
  
  # F5
  denom_beta <- (beta^2 * iter_prec) + iter_rec
  iter_fbeta <- if(denom_beta > 0) {
    ((1 + beta^2) * (iter_prec * iter_rec)) / denom_beta
  } else { 0 }
  
  dist_f1[i]   <- iter_f1
  dist_f5[i]   <- iter_fbeta
  iter_ActualNeg <- iter_TN + iter_FP
  iter_Total     <- iter_TP + iter_FP + iter_FN + iter_TN
  iter_spec      <- if(iter_ActualNeg > 0) iter_TN / iter_ActualNeg else 0

  # Larner's Metrics
  iter_Y    <- iter_rec + iter_spec - 1 # Youden Index
  iter_UAcc <- (iter_rec + iter_spec) / 2 # Unbiased Accuracy
  iter_Q    <- (iter_TP + iter_FP) / iter_Total # Bias / Positive Sign Rate
  iter_EI   <- (iter_TP + iter_TN - (iter_FP + iter_FN)) / iter_Total # Efficiency Index

  dist_Y[i]  <- iter_Y
  dist_EI[i] <- iter_EI
  dist_Q[i]  <- iter_Q

  # --- Calculation of Larner's Metrics (add inside the simulation loop) ---

  # Pre-calculate base probabilities for indices
  iter_Sens <- if(iter_ActualPos > 0) iter_TP / iter_ActualPos else 0
  iter_Spec <- if(iter_ActualNeg > 0) iter_TN / iter_ActualNeg else 0
  iter_PPV  <- if((iter_TP + iter_FP) > 0) iter_TP / (iter_TP + iter_FP) else 0
  iter_NPV  <- if((iter_TN + iter_FN) > 0) iter_TN / (iter_TN + iter_FN) else 0

  # 1. Critical Success Index (CSI) - Ignores True Negatives
  iter_CSI <- if((iter_TP + iter_FN + iter_FP) > 0) {
    iter_TP / (iter_TP + iter_FN + iter_FP) 
  } else { 0 }

  # 2. Predictive Summary Index (PSI) - Net gain of certainty
  iter_PSI <- iter_PPV + iter_NPV - 1

  # 3. Clinical Utility Indexes (PCUI & NCUI)
  iter_PCUI <- iter_Sens * iter_PPV
  iter_NCUI <- iter_Spec * iter_NPV

  # 4. Summary Utility Index (SUI) - Unitary score of utility 
  iter_SUI <- iter_PCUI + iter_NCUI

  dist_csi[i] <- iter_CSI
  dist_psi[i] <- iter_PSI
  dist_sui[i] <- iter_SUI
  dist_pcui[i] <- iter_PCUI
  dist_ncui[i] <- iter_NCUI
  dist_ActualPos[i] <- iter_ActualPos
  dist_ActualNeg[i] <- iter_ActualNeg
}

count_Above_AET <- total_TP + total_FN
count_Below_AET  <- total_TN + total_FP
FN_Rate_final <- total_FN / count_Above_AET
FP_Rate_final <- total_FP / count_Below_AET

recall <- total_TP / count_Above_AET
precision <- total_TP / (total_TP + total_FP)

F1 <- (2 * precision * recall / (precision + recall))
beta <- 5
f_beta <- (1 + beta^2) * (precision * recall) / ((beta^2 * precision) + recall)

SD_FN <- sqrt(FN_Rate_final * (1 - FN_Rate_final) / count_Above_AET)
SD_FP <- sqrt(FP_Rate_final * (1 - FP_Rate_final) / count_Below_AET)

summary_table <- bind_rows(
  summary_table,
  data.frame(
    Method = "Mean/Percentile (16th)",
    FN_Rate = FN_Rate_final,
    FP_Rate = FP_Rate_final,
    FN_SD = SD_FN,
    FP_SD = SD_FP,
    precision = precision,
    recall = recall,
    F1 = F1,
    f_beta = f_beta,
    Youden_Y = mean(dist_Y, na.rm = TRUE),
    Bias_Q = mean(dist_Q, na.rm = TRUE),
    Efficiency_EI = mean(dist_EI, na.rm = TRUE),
    MCC = mean(dist_MCC, na.rm = TRUE),
    DOR = mean(dist_DOR, na.rm = TRUE),
    
    # New Larner Metrics
    CSI = mean(dist_csi, na.rm = TRUE),
    PSI = mean(dist_psi, na.rm = TRUE),
    SUI = mean(dist_sui, na.rm = TRUE),
    PCUI = mean(dist_pcui, na.rm = TRUE),
    NCUI = mean(dist_ncui, na.rm = TRUE),

    # Absolute Counts and UF
    is_at_and_above_threshold = count_Above_AET,
    X.is_at_and_above_threshold = count_Below_AET,
    reported_at_and_above_threshold = total_TP + total_FP,
    X.reported_at_and_above_threshold = total_TN + total_FN,
    perc_is_at_and_above_threshold = (count_Above_AET / (count_Above_AET + count_Below_AET)) * 100,
    perc_reported_at_and_above_threshold = ((total_TP + total_FP) / (count_Above_AET + count_Below_AET)) * 100,
    UF = (meanLCMS / percentileLCMS)))

metrics_df <- data.frame(
  Method = "Mean/Percentile (16th)",
  F1 = dist_f1,
  F5 = dist_f5,
  Percent_FN = dist_fnr,
  Percent_FP = dist_fpr,
  EI = dist_EI,
  Youden_Y = dist_Y,
  Bias_Q = dist_Q,
  MCC = dist_MCC,
  DOR = dist_DOR
)

all_distributions <- rbind(all_distributions, metrics_df)

```

## LC-MS Correction RRFlow

This simulation assumes that we know the true RRF of all compounds sampled and therefore performing the conditional correction. We also force the UF to be positive by removing low and high responders.

```{r}

## RSD is calculated only for RRFs between 0.05 and 1 
## 'Compounds with RRF lower than 0.05 are excluded from UF calculation (20 times lower than the internal standard).page 8
## Therefore, as a criterion for RRF selection, only compounds with an RRF less than 1 are actually considered for UF calculation. page 8

rsd <- function(x) {
  x_use = x[x<=1&x>0.05]
  return(sd(x_use, na.rm = TRUE) / mean(x_use, na.rm = TRUE))
}

## Simulation starts
set.seed(123)
N_sim <- 10000
dist_Y <- numeric(N_sim)

dist_csi <- numeric(N_sim)
dist_psi <- numeric(N_sim)
dist_sui <- numeric(N_sim)
dist_pcui <- numeric(N_sim)
dist_ncui <- numeric(N_sim)
dist_ActualPos <- numeric(N_sim)
dist_ActualNeg <- numeric(N_sim)

dist_EI <- numeric(N_sim)
dist_Q <- numeric(N_sim)
dist_MCC <- numeric(N_sim)
dist_DOR <- numeric(N_sim)
N_extractables <- 80
total_FN <- 0
total_FP <- 0
total_TP <- 0
total_TN <- 0

# remove 0 values
RRF_LCMS_superior0 <- lcmsneg_clap_clean$"5 µg/mL"
RRF_LCMS_superior0 <- RRF_LCMS_superior0[RRF_LCMS_superior0 > 0]

# Apply Part 18 formula for database
UF <- 1/(1-rsd(rgamma(N_extractables, gamma_shape, gamma_rate)))

# Start simulation
for (i in 1:N_sim) {

## 1) Concentration around the AET
base_conc_sim <- runif(N_extractables, min = 0.5 * AET, max = 2 * AET)
## A) Sampling from the LC/MS- dataset fitted to gamma dist (5µg/mL – fitted distribution from CLAP list) (5µg/mL)
simulated_RRF <- rgamma(N_extractables, gamma_shape, gamma_rate)
## IV) RRflow condition
## add a normal / lognormal distribution to get a "exp" RRF with some variability e.g 30%
exp_RRF <- simulated_RRF * rnorm(n = N_extractables, 1, 0.3)
exp_conc <- base_conc_sim * simulated_RRF   # experimental concentration : True concentration with the actual RRF
corrected_conc <- exp_conc * UF             # corrected concentration: Exp * UF (conservative method)
condition <- (simulated_RRF < 0.5) | (simulated_RRF > 2) # RRFlow conditions rescaling factor should be focused only to those compounds which exhibits RRF< 0.5 and RRF>2 
corrected_conc[condition] <- exp_conc[condition] / exp_RRF[condition] # average value of the RRF RRFlow application in extractables amount rescaling ( The average RRF determined by the RRFlow will be used to re-calculate the amount of the corresponding compound detected in the extractables study , will assume that it is in the linearity range) 

  is_at_and_above_threshold <- base_conc_sim >= AET
  reported_at_and_above_threshold <- corrected_conc >= AET

## Count per iteration
  iter_TP <- sum(is_at_and_above_threshold & reported_at_and_above_threshold, na.rm = TRUE)
  iter_FP <- sum(!is_at_and_above_threshold & reported_at_and_above_threshold, na.rm = TRUE)
  iter_FN <- sum(is_at_and_above_threshold & !reported_at_and_above_threshold, na.rm = TRUE)
  iter_TN <- sum(!is_at_and_above_threshold & !reported_at_and_above_threshold, na.rm = TRUE)

  total_TP <- total_TP + iter_TP
  total_FP <- total_FP + iter_FP
  total_FN <- total_FN + iter_FN
  total_TN <- total_TN + iter_TN

  # Matthews Correlation Coefficient (MCC)
  denom_MCC <- sqrt(as.numeric(iter_TP+iter_FP)*(iter_TP+iter_FN)*(iter_TN+iter_FP)*(iter_TN+iter_FN))
  iter_MCC <- if(denom_MCC > 0) ((as.numeric(iter_TP)*iter_TN) - (as.numeric(iter_FP)*iter_FN)) / denom_MCC else 0
  
  # Diagnostic Odds Ratio (DOR)
  iter_DOR <- (as.numeric(iter_TP) * iter_TN) / ((as.numeric(iter_FP) * iter_FN) + 1e-6)
  
  dist_MCC[i] <- iter_MCC
  dist_DOR[i] <- iter_DOR


  # metrics per iterations
  iter_prec <- if((iter_TP + iter_FP) > 0) iter_TP / (iter_TP + iter_FP) else 0
  iter_rec  <- if((iter_TP + iter_FN) > 0) iter_TP / (iter_TP + iter_FN) else 0
  
  iter_ActualPos <- iter_TP + iter_FN
  iter_ActualNeg <- iter_TN + iter_FP

  iter_fnr <- if(iter_ActualPos > 0) iter_FN / iter_ActualPos else 0
  iter_fpr <- if(iter_ActualNeg > 0) iter_FP / iter_ActualNeg else 0
  
  dist_fnr[i] <- iter_fnr * 100
  dist_fpr[i] <- iter_fpr * 100
  
  # F1
  iter_f1 <- if((iter_prec + iter_rec) > 0) {
    (2 * iter_prec * iter_rec) / (iter_prec + iter_rec)
  } else { 0 }
  
  # F5
  denom_beta <- (beta^2 * iter_prec) + iter_rec
  iter_fbeta <- if(denom_beta > 0) {
    ((1 + beta^2) * (iter_prec * iter_rec)) / denom_beta
  } else { 0 }
  
  dist_f1[i]   <- iter_f1
  dist_f5[i]   <- iter_fbeta
  iter_ActualNeg <- iter_TN + iter_FP
  iter_Total     <- iter_TP + iter_FP + iter_FN + iter_TN
  iter_spec      <- if(iter_ActualNeg > 0) iter_TN / iter_ActualNeg else 0

  # Larner's Metrics
  iter_Y    <- iter_rec + iter_spec - 1 # Youden Index
  iter_UAcc <- (iter_rec + iter_spec) / 2 # Unbiased Accuracy
  iter_Q    <- (iter_TP + iter_FP) / iter_Total # Bias / Positive Sign Rate
  iter_EI   <- (iter_TP + iter_TN - (iter_FP + iter_FN)) / iter_Total # Efficiency Index

  dist_Y[i]  <- iter_Y
  dist_EI[i] <- iter_EI
  dist_Q[i]  <- iter_Q

  # --- Calculation of Larner's Metrics (add inside the simulation loop) ---

  # Pre-calculate base probabilities for indices
  iter_Sens <- if(iter_ActualPos > 0) iter_TP / iter_ActualPos else 0
  iter_Spec <- if(iter_ActualNeg > 0) iter_TN / iter_ActualNeg else 0
  iter_PPV  <- if((iter_TP + iter_FP) > 0) iter_TP / (iter_TP + iter_FP) else 0
  iter_NPV  <- if((iter_TN + iter_FN) > 0) iter_TN / (iter_TN + iter_FN) else 0

  # 1. Critical Success Index (CSI) - Ignores True Negatives
  iter_CSI <- if((iter_TP + iter_FN + iter_FP) > 0) {
    iter_TP / (iter_TP + iter_FN + iter_FP) 
  } else { 0 }

  # 2. Predictive Summary Index (PSI) - Net gain of certainty
  iter_PSI <- iter_PPV + iter_NPV - 1

  # 3. Clinical Utility Indexes (PCUI & NCUI)
  iter_PCUI <- iter_Sens * iter_PPV
  iter_NCUI <- iter_Spec * iter_NPV

  # 4. Summary Utility Index (SUI) - Unitary score of utility 
  iter_SUI <- iter_PCUI + iter_NCUI

  dist_csi[i] <- iter_CSI
  dist_psi[i] <- iter_PSI
  dist_sui[i] <- iter_SUI
  dist_pcui[i] <- iter_PCUI
  dist_ncui[i] <- iter_NCUI
  dist_ActualPos[i] <- iter_ActualPos
  dist_ActualNeg[i] <- iter_ActualNeg
}

count_Above_AET <- total_TP + total_FN
count_Below_AET  <- total_TN + total_FP
FN_Rate_final <- total_FN / count_Above_AET
FP_Rate_final <- total_FP / count_Below_AET

recall <- total_TP / count_Above_AET
precision <- total_TP / (total_TP + total_FP)

F1 <- (2 * precision * recall / (precision + recall))
beta <- 5
f_beta <- (1 + beta^2) * (precision * recall) / ((beta^2 * precision) + recall)

SD_FN <- sqrt(FN_Rate_final * (1 - FN_Rate_final) / count_Above_AET)
SD_FP <- sqrt(FP_Rate_final * (1 - FP_Rate_final) / count_Below_AET)

summary_table <- bind_rows(
  summary_table,
  data.frame(
    Method = "RRFlow",
    FN_Rate = FN_Rate_final,
    FP_Rate = FP_Rate_final,
    FN_SD = SD_FN,
    FP_SD = SD_FP,
    precision = precision,
    recall = recall,
    F1 = F1,
    f_beta = f_beta,
    Youden_Y = mean(dist_Y, na.rm = TRUE),
    Bias_Q = mean(dist_Q, na.rm = TRUE),
    Efficiency_EI = mean(dist_EI, na.rm = TRUE),
    MCC = mean(dist_MCC, na.rm = TRUE),
    DOR = mean(dist_DOR, na.rm = TRUE),
    
    # New Larner Metrics
    CSI = mean(dist_csi, na.rm = TRUE),
    PSI = mean(dist_psi, na.rm = TRUE),
    SUI = mean(dist_sui, na.rm = TRUE),
    PCUI = mean(dist_pcui, na.rm = TRUE),
    NCUI = mean(dist_ncui, na.rm = TRUE),

    # Absolute Counts and UF
    is_at_and_above_threshold = count_Above_AET,
    X.is_at_and_above_threshold = count_Below_AET,
    reported_at_and_above_threshold = total_TP + total_FP,
    X.reported_at_and_above_threshold = total_TN + total_FN,
    perc_is_at_and_above_threshold = (count_Above_AET / (count_Above_AET + count_Below_AET)) * 100,
    perc_reported_at_and_above_threshold = ((total_TP + total_FP) / (count_Above_AET + count_Below_AET)) * 100,
    UF = 1/(1-rsd(rgamma(N_extractables, gamma_shape, gamma_rate)))
  )
)

metrics_df <- data.frame(
  Method = "RRFlow",
  F1 = dist_f1,
  F5 = dist_f5,
  Percent_FN = dist_fnr,
  Percent_FP = dist_fpr,
  EI = dist_EI,
  Youden_Y = dist_Y,
  Bias_Q = dist_Q,
  MCC = dist_MCC,
  DOR = dist_DOR
)

all_distributions <- rbind(all_distributions, metrics_df)

```

## LC-MS Correction RRF\<1

This simulation assumes that we know the true RRF of all compounds sampled and therefore performing the conditional correction. We also force the UF to be positive by removing low and high responders.

```{r}

## RSD is calculated only for RRFs between 0.05 and 1 
## 'Compounds with RRF lower than 0.05 are excluded from UF calculation (20 times lower than the internal standard).page 8
## Therefore, as a criterion for RRF selection, only compounds with an RRF less than 1 are actually considered for UF calculation. page 8

rsd <- function(x) {
  x_use = x[x<=1&x>0.05]
  return(sd(x_use, na.rm = TRUE) / mean(x_use, na.rm = TRUE))
}

## Simulation starts
set.seed(123)
N_sim <- 10000
dist_Y <- numeric(N_sim)

dist_csi <- numeric(N_sim)
dist_psi <- numeric(N_sim)
dist_sui <- numeric(N_sim)
dist_pcui <- numeric(N_sim)
dist_ncui <- numeric(N_sim)
dist_ActualPos <- numeric(N_sim)
dist_ActualNeg <- numeric(N_sim)

dist_EI <- numeric(N_sim)
dist_Q <- numeric(N_sim)
dist_MCC <- numeric(N_sim)
dist_DOR <- numeric(N_sim)
N_extractables <- 80
total_FN <- 0
total_FP <- 0
total_TP <- 0
total_TN <- 0

# remove 0 values
RRF_LCMS_superior0 <- lcmsneg_clap_clean$"5 µg/mL"
RRF_LCMS_superior0 <- RRF_LCMS_superior0[RRF_LCMS_superior0 > 0]

# Apply Part 18 formula for database
UF <- 1/(1-rsd(rgamma(N_extractables, gamma_shape, gamma_rate)))

# Start simulation
for (i in 1:N_sim) {

## 1) Concentration around the AET
base_conc_sim <- runif(N_extractables, min = 0.5 * AET, max = 2 * AET)
## A) Sampling from the LC/MS- dataset fitted to gamma dist (5µg/mL – fitted distribution from CLAP list) (5µg/mL)
simulated_RRF <- rgamma(N_extractables, gamma_shape, gamma_rate)
## IV) RRflow condition
## add a normal / lognormal distribution to get a "exp" RRF with some variability e.g 30%
exp_RRF <- simulated_RRF * rnorm(n = N_extractables, 1, 0.3)
exp_conc <- base_conc_sim * simulated_RRF   # experimental concentration : True concentration with the actual RRF
corrected_conc <- exp_conc * UF             # corrected concentration: Exp * UF (conservative method)
condition <- (simulated_RRF < 1)  # Condition only for LOW responders
corrected_conc[condition] <- exp_conc[condition] / exp_RRF[condition] # average value of the RRF RRFlow application in extractables amount rescaling ( The average RRF determined by the RRFlow will be used to re-calculate the amount of the corresponding compound detected in the extractables study , will assume that it is in the linearity range) 

  is_at_and_above_threshold <- base_conc_sim >= AET
  reported_at_and_above_threshold <- corrected_conc >= AET

## Count per iteration
  iter_TP <- sum(is_at_and_above_threshold & reported_at_and_above_threshold, na.rm = TRUE)
  iter_FP <- sum(!is_at_and_above_threshold & reported_at_and_above_threshold, na.rm = TRUE)
  iter_FN <- sum(is_at_and_above_threshold & !reported_at_and_above_threshold, na.rm = TRUE)
  iter_TN <- sum(!is_at_and_above_threshold & !reported_at_and_above_threshold, na.rm = TRUE)

  total_TP <- total_TP + iter_TP
  total_FP <- total_FP + iter_FP
  total_FN <- total_FN + iter_FN
  total_TN <- total_TN + iter_TN

  # Matthews Correlation Coefficient (MCC)
  denom_MCC <- sqrt(as.numeric(iter_TP+iter_FP)*(iter_TP+iter_FN)*(iter_TN+iter_FP)*(iter_TN+iter_FN))
  iter_MCC <- if(denom_MCC > 0) ((as.numeric(iter_TP)*iter_TN) - (as.numeric(iter_FP)*iter_FN)) / denom_MCC else 0
  
  # Diagnostic Odds Ratio (DOR)
  iter_DOR <- (as.numeric(iter_TP) * iter_TN) / ((as.numeric(iter_FP) * iter_FN) + 1e-6)
  
  dist_MCC[i] <- iter_MCC
  dist_DOR[i] <- iter_DOR


  # metrics per iterations
  iter_prec <- if((iter_TP + iter_FP) > 0) iter_TP / (iter_TP + iter_FP) else 0
  iter_rec  <- if((iter_TP + iter_FN) > 0) iter_TP / (iter_TP + iter_FN) else 0
  
  iter_ActualPos <- iter_TP + iter_FN
  iter_ActualNeg <- iter_TN + iter_FP

  iter_fnr <- if(iter_ActualPos > 0) iter_FN / iter_ActualPos else 0
  iter_fpr <- if(iter_ActualNeg > 0) iter_FP / iter_ActualNeg else 0
  
  dist_fnr[i] <- iter_fnr * 100
  dist_fpr[i] <- iter_fpr * 100
  
  # F1
  iter_f1 <- if((iter_prec + iter_rec) > 0) {
    (2 * iter_prec * iter_rec) / (iter_prec + iter_rec)
  } else { 0 }
  
  # F5
  denom_beta <- (beta^2 * iter_prec) + iter_rec
  iter_fbeta <- if(denom_beta > 0) {
    ((1 + beta^2) * (iter_prec * iter_rec)) / denom_beta
  } else { 0 }
  
  dist_f1[i]   <- iter_f1
  dist_f5[i]   <- iter_fbeta
  iter_ActualNeg <- iter_TN + iter_FP
  iter_Total     <- iter_TP + iter_FP + iter_FN + iter_TN
  iter_spec      <- if(iter_ActualNeg > 0) iter_TN / iter_ActualNeg else 0

  # Larner's Metrics
  iter_Y    <- iter_rec + iter_spec - 1 # Youden Index
  iter_UAcc <- (iter_rec + iter_spec) / 2 # Unbiased Accuracy
  iter_Q    <- (iter_TP + iter_FP) / iter_Total # Bias / Positive Sign Rate
  iter_EI   <- (iter_TP + iter_TN - (iter_FP + iter_FN)) / iter_Total # Efficiency Index

  dist_Y[i]  <- iter_Y
  dist_EI[i] <- iter_EI
  dist_Q[i]  <- iter_Q

  # --- Calculation of Larner's Metrics (add inside the simulation loop) ---

  # Pre-calculate base probabilities for indices
  iter_Sens <- if(iter_ActualPos > 0) iter_TP / iter_ActualPos else 0
  iter_Spec <- if(iter_ActualNeg > 0) iter_TN / iter_ActualNeg else 0
  iter_PPV  <- if((iter_TP + iter_FP) > 0) iter_TP / (iter_TP + iter_FP) else 0
  iter_NPV  <- if((iter_TN + iter_FN) > 0) iter_TN / (iter_TN + iter_FN) else 0

  # 1. Critical Success Index (CSI) - Ignores True Negatives
  iter_CSI <- if((iter_TP + iter_FN + iter_FP) > 0) {
    iter_TP / (iter_TP + iter_FN + iter_FP) 
  } else { 0 }

  # 2. Predictive Summary Index (PSI) - Net gain of certainty
  iter_PSI <- iter_PPV + iter_NPV - 1

  # 3. Clinical Utility Indexes (PCUI & NCUI)
  iter_PCUI <- iter_Sens * iter_PPV
  iter_NCUI <- iter_Spec * iter_NPV

  # 4. Summary Utility Index (SUI) - Unitary score of utility 
  iter_SUI <- iter_PCUI + iter_NCUI

  dist_csi[i] <- iter_CSI
  dist_psi[i] <- iter_PSI
  dist_sui[i] <- iter_SUI
  dist_pcui[i] <- iter_PCUI
  dist_ncui[i] <- iter_NCUI
  dist_ActualPos[i] <- iter_ActualPos
  dist_ActualNeg[i] <- iter_ActualNeg
}

count_Above_AET <- total_TP + total_FN
count_Below_AET  <- total_TN + total_FP
FN_Rate_final <- total_FN / count_Above_AET
FP_Rate_final <- total_FP / count_Below_AET

recall <- total_TP / count_Above_AET
precision <- total_TP / (total_TP + total_FP)

F1 <- (2 * precision * recall / (precision + recall))
beta <- 5
f_beta <- (1 + beta^2) * (precision * recall) / ((beta^2 * precision) + recall)

SD_FN <- sqrt(FN_Rate_final * (1 - FN_Rate_final) / count_Above_AET)
SD_FP <- sqrt(FP_Rate_final * (1 - FP_Rate_final) / count_Below_AET)

summary_table <- bind_rows(
  summary_table,
  data.frame(
    Method = "RRF<1",
    FN_Rate = FN_Rate_final,
    FP_Rate = FP_Rate_final,
    FN_SD = SD_FN,
    FP_SD = SD_FP,
    precision = precision,
    recall = recall,
    F1 = F1,
    f_beta = f_beta,
    Youden_Y = mean(dist_Y, na.rm = TRUE),
    Bias_Q = mean(dist_Q, na.rm = TRUE),
    Efficiency_EI = mean(dist_EI, na.rm = TRUE),
    MCC = mean(dist_MCC, na.rm = TRUE),
    DOR = mean(dist_DOR, na.rm = TRUE),
    
    # New Larner Metrics
    CSI = mean(dist_csi, na.rm = TRUE),
    PSI = mean(dist_psi, na.rm = TRUE),
    SUI = mean(dist_sui, na.rm = TRUE),
    PCUI = mean(dist_pcui, na.rm = TRUE),
    NCUI = mean(dist_ncui, na.rm = TRUE),

    # Absolute Counts and UF
    is_at_and_above_threshold = count_Above_AET,
    X.is_at_and_above_threshold = count_Below_AET,
    reported_at_and_above_threshold = total_TP + total_FP,
    X.reported_at_and_above_threshold = total_TN + total_FN,
    perc_is_at_and_above_threshold = (count_Above_AET / (count_Above_AET + count_Below_AET)) * 100,
    perc_reported_at_and_above_threshold = ((total_TP + total_FP) / (count_Above_AET + count_Below_AET)) * 100,
    UF = 1/(1-rsd(rgamma(N_extractables, gamma_shape, gamma_rate)))
  )
)

metrics_df <- data.frame(
  Method = "RRF<1",
  F1 = dist_f1,
  F5 = dist_f5,
  Percent_FN = dist_fnr,
  Percent_FP = dist_fpr,
  EI = dist_EI,
  Youden_Y = dist_Y,
  Bias_Q = dist_Q,
  MCC = dist_MCC,
  DOR = dist_DOR
)

all_distributions <- rbind(all_distributions, metrics_df)

```

```{r}
UF
summary_table
```

## LC-MS Correction percentile 10th

Here we divide by the 10th percentile (which is arbitrary) assuming to capture the left hand side of the RRF database (low responders)

```{r}
set.seed(123)
N_sim <- 10000
dist_Y <- numeric(N_sim)

dist_csi <- numeric(N_sim)
dist_psi <- numeric(N_sim)
dist_sui <- numeric(N_sim)
dist_pcui <- numeric(N_sim)
dist_ncui <- numeric(N_sim)
dist_ActualPos <- numeric(N_sim)
dist_ActualNeg <- numeric(N_sim)

dist_EI <- numeric(N_sim)
dist_Q <- numeric(N_sim)
dist_MCC <- numeric(N_sim)
dist_DOR <- numeric(N_sim)
N_extractables <- 80
total_FN <- 0
total_FP <- 0
total_TP <- 0
total_TN <- 0

RRF_LCMS_superior0 <- lcmsneg_clap_clean$"5 µg/mL"
RRF_LCMS_superior0 <- RRF_LCMS_superior0[RRF_LCMS_superior0 > 0]

conservative_RRF_estimate <- quantile(rgamma(N_extractables, gamma_shape, gamma_rate), 0.10)

for (i in 1:N_sim) {

## 1) Concentration around the AET
base_conc_sim <- runif(N_extractables, min = 0.5 * AET, max = 2 * AET)
## A) Sampling from the LC/MS- dataset fitted to gamma dist (5µg/mL – fitted distribution from CLAP list) (5µg/mL)
simulated_RRF <- rgamma(N_extractables, gamma_shape, gamma_rate)
## V 

exp_conc <- base_conc_sim * simulated_RRF   # experimental concentration : True concentration with the actual RRF
corrected_conc <- exp_conc / conservative_RRF_estimate ## need to find rationale

is_at_and_above_threshold <- base_conc_sim >= AET
reported_at_and_above_threshold <- corrected_conc >= AET

## Count per iteration
  iter_TP <- sum(is_at_and_above_threshold & reported_at_and_above_threshold, na.rm = TRUE)
  iter_FP <- sum(!is_at_and_above_threshold & reported_at_and_above_threshold, na.rm = TRUE)
  iter_FN <- sum(is_at_and_above_threshold & !reported_at_and_above_threshold, na.rm = TRUE)
  iter_TN <- sum(!is_at_and_above_threshold & !reported_at_and_above_threshold, na.rm = TRUE)

  total_TP <- total_TP + iter_TP
  total_FP <- total_FP + iter_FP
  total_FN <- total_FN + iter_FN
  total_TN <- total_TN + iter_TN

  # Matthews Correlation Coefficient (MCC)
  denom_MCC <- sqrt(as.numeric(iter_TP+iter_FP)*(iter_TP+iter_FN)*(iter_TN+iter_FP)*(iter_TN+iter_FN))
  iter_MCC <- if(denom_MCC > 0) ((as.numeric(iter_TP)*iter_TN) - (as.numeric(iter_FP)*iter_FN)) / denom_MCC else 0
  
  # Diagnostic Odds Ratio (DOR)
  iter_DOR <- (as.numeric(iter_TP) * iter_TN) / ((as.numeric(iter_FP) * iter_FN) + 1e-6)
  
  dist_MCC[i] <- iter_MCC
  dist_DOR[i] <- iter_DOR


  # metrics per iterations
  iter_prec <- if((iter_TP + iter_FP) > 0) iter_TP / (iter_TP + iter_FP) else 0
  iter_rec  <- if((iter_TP + iter_FN) > 0) iter_TP / (iter_TP + iter_FN) else 0
  
  iter_ActualPos <- iter_TP + iter_FN
  iter_ActualNeg <- iter_TN + iter_FP

  iter_fnr <- if(iter_ActualPos > 0) iter_FN / iter_ActualPos else 0
  iter_fpr <- if(iter_ActualNeg > 0) iter_FP / iter_ActualNeg else 0
  
  dist_fnr[i] <- iter_fnr * 100
  dist_fpr[i] <- iter_fpr * 100
  
  # F1
  iter_f1 <- if((iter_prec + iter_rec) > 0) {
    (2 * iter_prec * iter_rec) / (iter_prec + iter_rec)
  } else { 0 }
  
  # F5
  denom_beta <- (beta^2 * iter_prec) + iter_rec
  iter_fbeta <- if(denom_beta > 0) {
    ((1 + beta^2) * (iter_prec * iter_rec)) / denom_beta
  } else { 0 }
  
  dist_f1[i]   <- iter_f1
  dist_f5[i]   <- iter_fbeta
  iter_ActualNeg <- iter_TN + iter_FP
  iter_Total     <- iter_TP + iter_FP + iter_FN + iter_TN
  iter_spec      <- if(iter_ActualNeg > 0) iter_TN / iter_ActualNeg else 0

  # Larner's Metrics
  iter_Y    <- iter_rec + iter_spec - 1 # Youden Index
  iter_UAcc <- (iter_rec + iter_spec) / 2 # Unbiased Accuracy
  iter_Q    <- (iter_TP + iter_FP) / iter_Total # Bias / Positive Sign Rate
  iter_EI   <- (iter_TP + iter_TN - (iter_FP + iter_FN)) / iter_Total # Efficiency Index

  dist_Y[i]  <- iter_Y
  dist_EI[i] <- iter_EI
  dist_Q[i]  <- iter_Q

  # --- Calculation of Larner's Metrics (add inside the simulation loop) ---

  # Pre-calculate base probabilities for indices
  iter_Sens <- if(iter_ActualPos > 0) iter_TP / iter_ActualPos else 0
  iter_Spec <- if(iter_ActualNeg > 0) iter_TN / iter_ActualNeg else 0
  iter_PPV  <- if((iter_TP + iter_FP) > 0) iter_TP / (iter_TP + iter_FP) else 0
  iter_NPV  <- if((iter_TN + iter_FN) > 0) iter_TN / (iter_TN + iter_FN) else 0

  # 1. Critical Success Index (CSI) - Ignores True Negatives
  iter_CSI <- if((iter_TP + iter_FN + iter_FP) > 0) {
    iter_TP / (iter_TP + iter_FN + iter_FP) 
  } else { 0 }

  # 2. Predictive Summary Index (PSI) - Net gain of certainty
  iter_PSI <- iter_PPV + iter_NPV - 1

  # 3. Clinical Utility Indexes (PCUI & NCUI)
  iter_PCUI <- iter_Sens * iter_PPV
  iter_NCUI <- iter_Spec * iter_NPV

  # 4. Summary Utility Index (SUI) - Unitary score of utility 
  iter_SUI <- iter_PCUI + iter_NCUI

  dist_csi[i] <- iter_CSI
  dist_psi[i] <- iter_PSI
  dist_sui[i] <- iter_SUI
  dist_pcui[i] <- iter_PCUI
  dist_ncui[i] <- iter_NCUI
  dist_ActualPos[i] <- iter_ActualPos
  dist_ActualNeg[i] <- iter_ActualNeg
}

count_Above_AET <- total_TP + total_FN
count_Below_AET  <- total_TN + total_FP
FN_Rate_final <- total_FN / count_Above_AET
FP_Rate_final <- total_FP / count_Below_AET

recall <- total_TP / count_Above_AET
precision <- total_TP / (total_TP + total_FP)

F1 <- (2 * precision * recall / (precision + recall))
beta <- 5
f_beta <- (1 + beta^2) * (precision * recall) / ((beta^2 * precision) + recall)

SD_FN <- sqrt(FN_Rate_final * (1 - FN_Rate_final) / count_Above_AET)
SD_FP <- sqrt(FP_Rate_final * (1 - FP_Rate_final) / count_Below_AET)

summary_table <- bind_rows(
  summary_table,
  data.frame(
    Method = "Percentile 10th",
    FN_Rate = FN_Rate_final,
    FP_Rate = FP_Rate_final,
    FN_SD = SD_FN,
    FP_SD = SD_FP,
    precision = precision,
    recall = recall,
    F1 = F1,
    f_beta = f_beta,
    Youden_Y = mean(dist_Y, na.rm = TRUE),
    Bias_Q = mean(dist_Q, na.rm = TRUE),
    Efficiency_EI = mean(dist_EI, na.rm = TRUE),
    MCC = mean(dist_MCC, na.rm = TRUE),
    DOR = mean(dist_DOR, na.rm = TRUE),
    
    # New Larner Metrics
    CSI = mean(dist_csi, na.rm = TRUE),
    PSI = mean(dist_psi, na.rm = TRUE),
    SUI = mean(dist_sui, na.rm = TRUE),
    PCUI = mean(dist_pcui, na.rm = TRUE),
    NCUI = mean(dist_ncui, na.rm = TRUE),

    # Absolute Counts and UF
    is_at_and_above_threshold = count_Above_AET,
    X.is_at_and_above_threshold = count_Below_AET,
    reported_at_and_above_threshold = total_TP + total_FP,
    X.reported_at_and_above_threshold = total_TN + total_FN,
    perc_is_at_and_above_threshold = (count_Above_AET / (count_Above_AET + count_Below_AET)) * 100,
    perc_reported_at_and_above_threshold = ((total_TP + total_FP) / (count_Above_AET + count_Below_AET)) * 100,
    UF = 1 / conservative_RRF_estimate
  )
)

metrics_df <- data.frame(
  Method = "Percentile 10th",
  F1 = dist_f1,
  F5 = dist_f5,
  Percent_FN = dist_fnr,
  Percent_FP = dist_fpr,
  EI = dist_EI,
  Youden_Y = dist_Y,
  Bias_Q = dist_Q,
  MCC = dist_MCC,
  DOR = dist_DOR
)

all_distributions <- rbind(all_distributions, metrics_df)

```

## LC-MS Correction percentile 20th

Here we divide by the 20th percentile (which is arbitrary) assuming to capture the left hand side of the RRF database (low responders)

```{r}
set.seed(123)
N_sim <- 10000
dist_Y <- numeric(N_sim)

dist_csi <- numeric(N_sim)
dist_psi <- numeric(N_sim)
dist_sui <- numeric(N_sim)
dist_pcui <- numeric(N_sim)
dist_ncui <- numeric(N_sim)
dist_ActualPos <- numeric(N_sim)
dist_ActualNeg <- numeric(N_sim)

dist_EI <- numeric(N_sim)
dist_Q <- numeric(N_sim)
dist_MCC <- numeric(N_sim)
dist_DOR <- numeric(N_sim)
N_extractables <- 80
total_FN <- 0
total_FP <- 0
total_TP <- 0
total_TN <- 0

RRF_LCMS_superior0 <- lcmsneg_clap_clean$"5 µg/mL"
RRF_LCMS_superior0 <- RRF_LCMS_superior0[RRF_LCMS_superior0 > 0]

conservative_RRF_estimate <- quantile(rgamma(N_extractables, gamma_shape, gamma_rate), 0.20)

for (i in 1:N_sim) {

## 1) Concentration around the AET
base_conc_sim <- runif(N_extractables, min = 0.5 * AET, max = 2 * AET)
## A) Sampling from the LC/MS- dataset fitted to gamma dist (5µg/mL – fitted distribution from CLAP list) (5µg/mL)
simulated_RRF <- rgamma(N_extractables, gamma_shape, gamma_rate)
## V 

exp_conc <- base_conc_sim * simulated_RRF   # experimental concentration : True concentration with the actual RRF
corrected_conc <- exp_conc / conservative_RRF_estimate ## need to find rationale

is_at_and_above_threshold <- base_conc_sim >= AET
reported_at_and_above_threshold <- corrected_conc >= AET
## Count per iteration
  iter_TP <- sum(is_at_and_above_threshold & reported_at_and_above_threshold, na.rm = TRUE)
  iter_FP <- sum(!is_at_and_above_threshold & reported_at_and_above_threshold, na.rm = TRUE)
  iter_FN <- sum(is_at_and_above_threshold & !reported_at_and_above_threshold, na.rm = TRUE)
  iter_TN <- sum(!is_at_and_above_threshold & !reported_at_and_above_threshold, na.rm = TRUE)

  total_TP <- total_TP + iter_TP
  total_FP <- total_FP + iter_FP
  total_FN <- total_FN + iter_FN
  total_TN <- total_TN + iter_TN

  # Matthews Correlation Coefficient (MCC)
  denom_MCC <- sqrt(as.numeric(iter_TP+iter_FP)*(iter_TP+iter_FN)*(iter_TN+iter_FP)*(iter_TN+iter_FN))
  iter_MCC <- if(denom_MCC > 0) ((as.numeric(iter_TP)*iter_TN) - (as.numeric(iter_FP)*iter_FN)) / denom_MCC else 0
  
  # Diagnostic Odds Ratio (DOR)
  iter_DOR <- (as.numeric(iter_TP) * iter_TN) / ((as.numeric(iter_FP) * iter_FN) + 1e-6)
  
  dist_MCC[i] <- iter_MCC
  dist_DOR[i] <- iter_DOR


  # metrics per iterations
  iter_prec <- if((iter_TP + iter_FP) > 0) iter_TP / (iter_TP + iter_FP) else 0
  iter_rec  <- if((iter_TP + iter_FN) > 0) iter_TP / (iter_TP + iter_FN) else 0
  
  iter_ActualPos <- iter_TP + iter_FN
  iter_ActualNeg <- iter_TN + iter_FP

  iter_fnr <- if(iter_ActualPos > 0) iter_FN / iter_ActualPos else 0
  iter_fpr <- if(iter_ActualNeg > 0) iter_FP / iter_ActualNeg else 0
  
  dist_fnr[i] <- iter_fnr * 100
  dist_fpr[i] <- iter_fpr * 100
  
  # F1
  iter_f1 <- if((iter_prec + iter_rec) > 0) {
    (2 * iter_prec * iter_rec) / (iter_prec + iter_rec)
  } else { 0 }
  
  # F5
  denom_beta <- (beta^2 * iter_prec) + iter_rec
  iter_fbeta <- if(denom_beta > 0) {
    ((1 + beta^2) * (iter_prec * iter_rec)) / denom_beta
  } else { 0 }
  
  dist_f1[i]   <- iter_f1
  dist_f5[i]   <- iter_fbeta
  iter_ActualNeg <- iter_TN + iter_FP
  iter_Total     <- iter_TP + iter_FP + iter_FN + iter_TN
  iter_spec      <- if(iter_ActualNeg > 0) iter_TN / iter_ActualNeg else 0

  # Larner's Metrics
  iter_Y    <- iter_rec + iter_spec - 1 # Youden Index
  iter_UAcc <- (iter_rec + iter_spec) / 2 # Unbiased Accuracy
  iter_Q    <- (iter_TP + iter_FP) / iter_Total # Bias / Positive Sign Rate
  iter_EI   <- (iter_TP + iter_TN - (iter_FP + iter_FN)) / iter_Total # Efficiency Index

  dist_Y[i]  <- iter_Y
  dist_EI[i] <- iter_EI
  dist_Q[i]  <- iter_Q

  # --- Calculation of Larner's Metrics (add inside the simulation loop) ---

  # Pre-calculate base probabilities for indices
  iter_Sens <- if(iter_ActualPos > 0) iter_TP / iter_ActualPos else 0
  iter_Spec <- if(iter_ActualNeg > 0) iter_TN / iter_ActualNeg else 0
  iter_PPV  <- if((iter_TP + iter_FP) > 0) iter_TP / (iter_TP + iter_FP) else 0
  iter_NPV  <- if((iter_TN + iter_FN) > 0) iter_TN / (iter_TN + iter_FN) else 0

  # 1. Critical Success Index (CSI) - Ignores True Negatives
  iter_CSI <- if((iter_TP + iter_FN + iter_FP) > 0) {
    iter_TP / (iter_TP + iter_FN + iter_FP) 
  } else { 0 }

  # 2. Predictive Summary Index (PSI) - Net gain of certainty
  iter_PSI <- iter_PPV + iter_NPV - 1

  # 3. Clinical Utility Indexes (PCUI & NCUI)
  iter_PCUI <- iter_Sens * iter_PPV
  iter_NCUI <- iter_Spec * iter_NPV

  # 4. Summary Utility Index (SUI) - Unitary score of utility 
  iter_SUI <- iter_PCUI + iter_NCUI

  dist_csi[i] <- iter_CSI
  dist_psi[i] <- iter_PSI
  dist_sui[i] <- iter_SUI
  dist_pcui[i] <- iter_PCUI
  dist_ncui[i] <- iter_NCUI
  dist_ActualPos[i] <- iter_ActualPos
  dist_ActualNeg[i] <- iter_ActualNeg
}

count_Above_AET <- total_TP + total_FN
count_Below_AET  <- total_TN + total_FP
FN_Rate_final <- total_FN / count_Above_AET
FP_Rate_final <- total_FP / count_Below_AET

recall <- total_TP / count_Above_AET
precision <- total_TP / (total_TP + total_FP)

F1 <- (2 * precision * recall / (precision + recall))
beta <- 5
f_beta <- (1 + beta^2) * (precision * recall) / ((beta^2 * precision) + recall)

SD_FN <- sqrt(FN_Rate_final * (1 - FN_Rate_final) / count_Above_AET)
SD_FP <- sqrt(FP_Rate_final * (1 - FP_Rate_final) / count_Below_AET)

summary_table <- bind_rows(
  summary_table,
  data.frame(
    Method = "Percentile 20th",
    FN_Rate = FN_Rate_final,
    FP_Rate = FP_Rate_final,
    FN_SD = SD_FN,
    FP_SD = SD_FP,
    precision = precision,
    recall = recall,
    F1 = F1,
    f_beta = f_beta,
    Youden_Y = mean(dist_Y, na.rm = TRUE),
    Bias_Q = mean(dist_Q, na.rm = TRUE),
    Efficiency_EI = mean(dist_EI, na.rm = TRUE),
    MCC = mean(dist_MCC, na.rm = TRUE),
    DOR = mean(dist_DOR, na.rm = TRUE),
    
    # New Larner Metrics
    CSI = mean(dist_csi, na.rm = TRUE),
    PSI = mean(dist_psi, na.rm = TRUE),
    SUI = mean(dist_sui, na.rm = TRUE),
    PCUI = mean(dist_pcui, na.rm = TRUE),
    NCUI = mean(dist_ncui, na.rm = TRUE),

    # Absolute Counts and UF
    is_at_and_above_threshold = count_Above_AET,
    X.is_at_and_above_threshold = count_Below_AET,
    reported_at_and_above_threshold = total_TP + total_FP,
    X.reported_at_and_above_threshold = total_TN + total_FN,
    perc_is_at_and_above_threshold = (count_Above_AET / (count_Above_AET + count_Below_AET)) * 100,
    perc_reported_at_and_above_threshold = ((total_TP + total_FP) / (count_Above_AET + count_Below_AET)) * 100,
    UF = 1 / conservative_RRF_estimate
  )
)

metrics_df <- data.frame(
  Method = "Percentile 20th",
  F1 = dist_f1,
  F5 = dist_f5,
  Percent_FN = dist_fnr,
  Percent_FP = dist_fpr,
  EI = dist_EI,
  Youden_Y = dist_Y,
  Bias_Q = dist_Q,
  MCC = dist_MCC,
  DOR = dist_DOR
)

all_distributions <- rbind(all_distributions, metrics_df)

```

```{r}
summary_table
```

```{r}
print(summary_table)
```

```{r}
summary_table
```

# Comparisons

```{r}
plot_data <- summary_table %>%
  pivot_longer(cols = c(FN_Rate, FP_Rate), names_to = "Type", values_to = "Rate") %>%
  mutate(
    SD = ifelse(Type == "FN_Rate", round(FN_SD*100, 2), round(FP_SD*100, 2)),
    Type = ifelse(Type == "FN_Rate", "False Negative", "False Positive")
  )

method_levels <- c(
  "No Correction",
  "UF (2)", 
  "UF (4)", 
  "UF (5)", 
  "UF (10)",  
  "UF (RSD)",
  "Mean/Percentile (16th)", 
  "RRFlow",
  "RRF<1",
  "Percentile 10th",
  "Percentile 20th")

plot_data$Method <- factor(plot_data$Method, levels = method_levels)


ggplot(plot_data, aes(x = Method, y = Rate*100, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), alpha = 0.8) +
  geom_errorbar(aes(ymin = pmax(0, Rate*100 - 1.96*SD), ymax = Rate*100 + 1.96*SD), 
                position = position_dodge(width = 0.9), width = 0.2) +
  labs(
    title = paste0("Comparison of LC-MS- Correction Methods (5 µg/mL)", " AET=", AET, "µg/mL"),
    subtitle = "Conc. Uniform distribution (around AET), RRF sampled from fitted dist fitted with gamma dist",
    y = "Rate (%)",
    x = "Method") +
  geom_text(aes(label = sprintf("%.1f%%", Rate*100)), position = position_dodge(width = 0.9), hjust = -0.5, size = 3, vjust = 0.1) +
  scale_fill_manual(values = c("False Negative" = "red", "False Positive" = "grey")) + 
  scale_y_continuous(limits = c(0, 100)) +
  theme_minimal() +
  coord_flip()

```

# Comparison metrics F5

```{r}
all_distributions$Method <- factor(all_distributions$Method, levels = method_levels)

ggplot(all_distributions, aes(x = Method, y = F5)) +
  geom_violin(fill = "orange", alpha = 0.6) +
  geom_boxplot(width = 0.1, color = "black", alpha = 0.5) +
    labs(
      title = paste0("Distribution of F5 scores | ", paste0("Comparison of LC-MS- Correction Methods (5 µg/mL)", " AET=", AET, "µg/mL")),
      subtitle = "Conc. Uniform distribution (around AET), RRF sampled from fitted dist") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Subset summary_table to the five threshold/UF columns and render as PNG
threshold_metrics_LC_MS__5_1A <- summary_table %>%
  dplyr::select(
    Method,
    is_at_and_above_threshold,
    X.is_at_and_above_threshold,
    reported_at_and_above_threshold,
    X.reported_at_and_above_threshold,
    perc_is_at_and_above_threshold,
    perc_reported_at_and_above_threshold,
    UF
  )

# Save the table as a PNG using gt
threshold_metrics_LC_MS__5_1A
```

# FP vs FN

```{r}
plot_data_FN_FP <- pivot_longer(all_distributions, 
                          cols = c("F5", "Percent_FN", "Percent_FP"), 
                          names_to = "Metric", 
                          values_to = "Value")

plot_data_FN_FP$Method <- factor(plot_data_FN_FP$Method, levels = method_levels)

```

```{r}
cluster_centers <- all_distributions %>%
  group_by(Method) %>%
  summarize(
    center_FP = mean(Percent_FP, na.rm = TRUE),
    center_FN = mean(Percent_FN, na.rm = TRUE)
  )

ggplot(all_distributions, aes(x = Percent_FP, y = Percent_FN, color = Method)) +
  # Plot the individual simulation dots
  geom_point(alpha = 0.05, size = 1) +
  
  # Draw the 95% confidence ellipses
  stat_ellipse(linewidth = 1, linetype = 1, type = "norm", level = 0.95) +
  
  geom_label_repel(
    data = cluster_centers, 
    aes(x = center_FP, y = center_FN, label = Method), 
    size = 4,
    fontface = "bold",
    alpha = 0.9,
    show.legend = FALSE,
    box.padding = 0.5,         # Adds "repel force" around the label itself
    point.padding = 0.5,       # Keeps the label from sitting directly on the exact center coordinate
    segment.color = "black",   # Makes the pointer line black so it stands out
    segment.size = 0.8,        # Makes the pointer line a bit thicker
    max.overlaps = Inf
  ) +
  
  # Clean up the labels and aesthetics
        labs(
      title = "Trade-off: False Positives vs. False Negatives",
      subtitle = paste0("10,000 Simulations per Method | AET = ", AET, " µg/mL"),
      x = "False Positive Rate (%)",
      y = "False Negative Rate (%)",
      color = "Correction Method"
    ) +
  theme_minimal() +
  theme(
    # change "bottom" to "none"!
    legend.position = "none", 
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold")
  )
```


# Matthews Correlation Coefficient and Diagnostic Odds Ratio (Larner 2024)

```{r}
# 1. MCC Boxplot (Comparison of Reliability)
ggplot(all_distributions, aes(x = Method, y = MCC, fill = Method)) +
  geom_boxplot(outlier.alpha = 0.1, color = "black") + 
  stat_summary(fun = mean, geom = "point", shape = 23, size = 3, fill = "white") +
  labs(
    title = "Matthews Correlation Coefficient (MCC) by Method",
    subtitle = "Higher values indicate better overall classification reliability (Larner 2024)",
    x = "Correction Method",
    y = "MCC (Reliability Score)"
  ) +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))

# 2. MCC Density Plot (Distribution Overlap)
ggplot(all_distributions, aes(x = MCC, fill = Method)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Distribution of MCC Scores",
    subtitle = "Comparison of classification reliability across 10,000 simulations",
    x = "Matthews Correlation Coefficient (MCC)",
    y = "Density"
  ) +
  theme_minimal()

# 3. DOR vs. MCC Scatter Plot (Utilizing ggrepel)
# Create centroids for labeling
method_centroids_larner <- all_distributions %>%
  group_by(Method) %>%
  summarize(mean_MCC = mean(MCC, na.rm = TRUE), 
            mean_DOR = mean(DOR, na.rm = TRUE))

ggplot(all_distributions, aes(x = log10(DOR + 1), y = MCC, color = Method)) +
  geom_point(alpha = 0.05, size = 0.5) +
  geom_point(data = method_centroids_larner, aes(x = log10(mean_DOR + 1), y = mean_MCC), 
             size = 4, shape = 21, fill = "white", stroke = 1.5) +
  geom_label_repel(data = method_centroids_larner, 
                   aes(x = log10(mean_DOR + 1), y = mean_MCC, label = Method),
                   size = 3, fontface = "bold", max.overlaps = Inf) +
  labs(
    title = "Discriminative Power (DOR) vs. Reliability (MCC)",
    subtitle = "DOR is shown on a Log10 scale | High values in both axes represent superior RRF correction",
    x = "Log10(Diagnostic Odds Ratio)",
    y = "Matthews Correlation Coefficient"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

# Confusion Matrices (Average over 10,000 Simulations)

```{r}
# ── Reconstruct TP / FP / FN / TN from the cumulated counts in summary_table ──
# Column meanings (sums over N_sim iterations):
#   is_at_and_above_threshold        = total_TP + total_FN  (all actual positives)
#   X.is_at_and_above_threshold      = total_TN + total_FP  (all actual negatives)
#   reported_at_and_above_threshold  = total_TP + total_FP  (all predicted positives)
#   X.reported_at_and_above_threshold= total_TN + total_FN  (all predicted negatives)

cm_data <- summary_table %>%
  mutate(
    total_grand = is_at_and_above_threshold + X.is_at_and_above_threshold,
    # Derive the four cells
    TP = recall * is_at_and_above_threshold,          # recall = TP / ActualPos
    FN = is_at_and_above_threshold - TP,
    FP = reported_at_and_above_threshold - TP,
    TN = X.is_at_and_above_threshold - FP,
    # Convert to % of grand total per method
    pct_TP = round(TP / total_grand * 100, 1),
    pct_FP = round(FP / total_grand * 100, 1),
    pct_FN = round(FN / total_grand * 100, 1),
    pct_TN = round(TN / total_grand * 100, 1)
  ) %>%
  dplyr::select(Method, pct_TP, pct_FP, pct_FN, pct_TN)

# ── Compute per-iteration SD from all_distributions ──
# Percent_FN = FN / ActualPos * 100  → actual_pos fraction = is_at_and_above_threshold / total_grand
# Percent_FP = FP / ActualNeg * 100  → actual_neg fraction = X.is_at_and_above_threshold / total_grand
cm_sd <- all_distributions %>%
  left_join(
    summary_table %>%
      transmute(
        Method,
        total_grand_per_sim = (is_at_and_above_threshold + X.is_at_and_above_threshold) / 10000,
        actual_pos_frac     = is_at_and_above_threshold / (is_at_and_above_threshold + X.is_at_and_above_threshold),
        actual_neg_frac     = X.is_at_and_above_threshold / (is_at_and_above_threshold + X.is_at_and_above_threshold)
      ),
    by = "Method"
  ) %>%
  mutate(
    sim_pct_FN = (Percent_FN / 100) * actual_pos_frac * 100,
    sim_pct_FP = (Percent_FP / 100) * actual_neg_frac * 100,
    sim_pct_TP = actual_pos_frac * 100 - sim_pct_FN,
    sim_pct_TN = actual_neg_frac * 100 - sim_pct_FP
  ) %>%
  group_by(Method) %>%
  summarize(
    sd_TP = round(sd(sim_pct_TP, na.rm = TRUE), 1),
    sd_FP = round(sd(sim_pct_FP, na.rm = TRUE), 1),
    sd_FN = round(sd(sim_pct_FN, na.rm = TRUE), 1),
    sd_TN = round(sd(sim_pct_TN, na.rm = TRUE), 1)
  )

# ── Pivot to long format for ggplot ──
cm_long <- cm_data %>%
  pivot_longer(
    cols = c(pct_TP, pct_FP, pct_FN, pct_TN),
    names_to  = "Cell",
    values_to = "Pct"
  ) %>%
  left_join(
    cm_sd %>%
      pivot_longer(
        cols = c(sd_TP, sd_FP, sd_FN, sd_TN),
        names_to  = "Cell_sd",
        values_to = "SD"
      ) %>%
      mutate(Cell = sub("sd_", "pct_", Cell_sd)) %>%
      dplyr::select(Method, Cell, SD),
    by = c("Method", "Cell")
  ) %>%
  mutate(
    Predicted = case_when(
      Cell %in% c("pct_TP", "pct_FP") ~ "Predicted Positive",
      TRUE                             ~ "Predicted Negative"
    ),
    Actual = case_when(
      Cell %in% c("pct_TP", "pct_FN") ~ "True Positive",
      TRUE                             ~ "True Negative"
    ),
    # Colour helper: correct cells vs error cells
    CellType = case_when(
      Cell == "pct_TP" ~ "TP",
      Cell == "pct_TN" ~ "TN",
      Cell == "pct_FP" ~ "FP",
      Cell == "pct_FN" ~ "FN"
    ),
    Label = paste0(Pct, "%\n\u00b1", SD, "%")
  ) %>%
  mutate(
    Predicted = factor(Predicted,
      levels = c("Predicted Positive",
                 "Predicted Negative")),
    Actual = factor(Actual,
      levels = c("True Positive",
                 "True Negative"))
  )

cm_long$Method <- factor(cm_long$Method, levels = method_levels)

# ── Plot ──
ggplot(cm_long, aes(x = Predicted, y = Actual, fill = CellType)) +
  geom_tile(color = "white", linewidth = 1.2) +
  geom_text(aes(label = Label), size = 3.8, fontface = "bold", color = "white", lineheight = 0.9) +
  scale_fill_manual(
    values = c(
      "TP" = "#2E86AB",   # blue  – correct positive
      "TN" = "#2E86AB",   # blue  – correct negative
      "FN" = "#E84855",   # red   – false alarm
      "FP" = "#F4A261"    # orange – missed
    ),
    guide = guide_legend(title = "Cell type")
  ) +
  facet_wrap(~ Method, ncol = 3) +
  labs(
    title    = "Average Confusion Matrix per Correction Method",
    subtitle = paste0("Values are mean \u00b1 SD (%) of all ", format(10000 * 80, big.mark = ","),
                      " compound–simulation pairs | AET = ", AET, " µg/mL"),
    x = "Predicted class (reported concentration vs AET)",
    y = "Actual class (true concentration vs AET)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    strip.text       = element_text(face = "bold", size = 10),
    axis.text        = element_text(size = 9),
    panel.grid       = element_blank(),
    legend.position  = "bottom",
    plot.title       = element_text(face = "bold", size = 13),
    plot.subtitle    = element_text(size = 10, color = "grey40")
  )

cm_data
```

```{r}
# Define number of simulations
N_sim <- 10000

# ── Reconstruct TP / FP / FN / TN and calculate averages per simulation ──
cm_data <- summary_table %>%
  mutate(
    # Derive the four total cells across all simulations
    TP = recall * is_at_and_above_threshold,          # recall = TP / ActualPos
    FN = is_at_and_above_threshold - TP,
    FP = reported_at_and_above_threshold - TP,
    TN = X.is_at_and_above_threshold - FP,

    # Convert to average count per simulation (out of 80)
    # Since total instances = 80 * 10,000, dividing by 10,000 gives count out of 80
    avg_TP = round(TP / N_sim),
    avg_FP = round(FP / N_sim),
    avg_FN = round(FN / N_sim),
    avg_TN = round(TN / N_sim)
  ) %>%
  dplyr::select(Method, avg_TP, avg_FP, avg_FN, avg_TN)

# ── Pivot to long format for ggplot ──
cm_long <- cm_data %>%
  pivot_longer(
    cols = c(avg_TP, avg_FP, avg_FN, avg_TN),
    names_to  = "Cell",
    values_to = "Count"
  ) %>%
  mutate(
    Predicted = case_when(
      Cell %in% c("avg_TP", "avg_FP") ~ "Predicted Positive",
      TRUE                           ~ "Predicted Negative"
    ),
    Actual = case_when(
      Cell %in% c("avg_TP", "avg_FN") ~ "True Positive",
      TRUE                           ~ "True Negative"
    ),
    CellType = case_when(
      Cell == "avg_TP" ~ "TP",
      Cell == "avg_TN" ~ "TN",
      Cell == "avg_FP" ~ "FP",
      Cell == "avg_FN" ~ "FN"
    ),
    Label = as.character(Count) # Removed the "%" sign
  ) %>%
  mutate(
    Predicted = factor(Predicted, levels = c("Predicted Positive", "Predicted Negative")),
    Actual = factor(Actual, levels = c("True Positive", "True Negative"))
  )

cm_long$Method <- factor(cm_long$Method, levels = method_levels)

# ── Plot ──
ggplot(cm_long, aes(x = Predicted, y = Actual, fill = CellType)) +
  geom_tile(color = "white", linewidth = 1.2) +
  geom_text(aes(label = Label), size = 4.5, fontface = "bold", color = "white") +
  scale_fill_manual(
    values = c(
      "TP" = "#2E86AB",   # blue  – correct positive
      "TN" = "#2E86AB",   # blue  – correct negative
      "FN" = "#E84855",   # red   – false alarm
      "FP" = "#F4A261"    # orange – missed
    ),
    guide = guide_legend(title = "Cell type")
  ) +
  facet_wrap(~ Method, ncol = 3) +
  labs(
    title    = "Average Confusion Matrix per Simulation",
    subtitle = paste0("Values are average counts out of 80 compounds | AET = ", AET, " µg/mL"),
    x = "Predicted class (reported concentration vs AET)",
    y = "Actual class (true concentration vs AET)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    strip.text       = element_text(face = "bold", size = 10),
    axis.text        = element_text(size = 9),
    panel.grid       = element_blank(),
    legend.position  = "bottom",
    plot.title       = element_text(face = "bold", size = 13),
    plot.subtitle    = element_text(size = 10, color = "grey40")
  )

cm_data
```

```{r}
# Define number of simulations
N_sim <- 10000

# ── Reconstruct TP / FP / FN / TN and calculate averages per simulation ──
cm_data_counts <- summary_table %>%
  mutate(
    TP = recall * is_at_and_above_threshold,
    FN = is_at_and_above_threshold - TP,
    FP = reported_at_and_above_threshold - TP,
    TN = X.is_at_and_above_threshold - FP,
    avg_TP = round(TP / N_sim),
    avg_FP = round(FP / N_sim),
    avg_FN = round(FN / N_sim),
    avg_TN = round(TN / N_sim)
  ) %>%
  dplyr::select(Method, avg_TP, avg_FP, avg_FN, avg_TN)

# ── Pivot to long format for ggplot ──
cm_long_counts <- cm_data_counts %>%
  pivot_longer(
    cols = c(avg_TP, avg_FP, avg_FN, avg_TN),
    names_to  = "Cell",
    values_to = "Count"
  ) %>%
  mutate(
    Predicted = case_when(
      Cell %in% c("avg_TP", "avg_FP") ~ "Pred. Positive",
      TRUE                           ~ "Pred. Negative"
    ),
    Actual = case_when(
      Cell %in% c("avg_TP", "avg_FN") ~ "True Positive",
      TRUE                           ~ "True Negative"
    ),
    CellType = case_when(
      Cell == "avg_TP" ~ "TP",
      Cell == "avg_TN" ~ "TN",
      Cell == "avg_FP" ~ "FP",
      Cell == "avg_FN" ~ "FN"
    ),
    Label = as.character(Count)
  ) %>%
  mutate(
    Predicted = factor(Predicted, levels = c("Pred. Positive", "Pred. Negative")),
    Actual = factor(Actual, levels = c("True Positive", "True Negative"))
  )

cm_long_counts$Method <- factor(cm_long_counts$Method, levels = method_levels)

# ── Plot ──
ggplot(cm_long_counts, aes(x = Predicted, y = Actual, fill = CellType)) +
  geom_tile(color = "white", linewidth = 1.2) +
  geom_text(aes(label = Label), size = 4.5, fontface = "bold", color = "white") +
  scale_fill_manual(
    values = c(
      "TP" = "#2E86AB",
      "TN" = "#2E86AB",
      "FN" = "#E84855",
      "FP" = "#F4A261"
    ),
    guide = guide_legend(title = "Cell type")
  ) +
  facet_wrap(~ Method, ncol = 3) +
  labs(
    title    = "Average Confusion Matrix per Simulation",
    subtitle = paste0("Values are average counts out of 80 compounds | AET = ", AET, " µg/mL"),
    x = "Predicted class (reported concentration vs AET)",
    y = "Actual class (true concentration vs AET)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    strip.text       = element_text(face = "bold", size = 10),
    axis.text        = element_text(size = 9),
    panel.grid       = element_blank(),
    legend.position  = "bottom",
    plot.title       = element_text(face = "bold", size = 13),
    plot.subtitle    = element_text(size = 10, color = "grey40")
  )

cm_data_counts
```

```{r}
# Average count per simulation WITH ± SD
cm_sd_counts <- all_distributions %>%
  left_join(
    summary_table %>%
      transmute(
        Method,
        actual_pos_per_sim = is_at_and_above_threshold / N_sim,
        actual_neg_per_sim = X.is_at_and_above_threshold / N_sim
      ),
    by = "Method"
  ) %>%
  mutate(
    sim_FN = (Percent_FN / 100) * actual_pos_per_sim,
    sim_FP = (Percent_FP / 100) * actual_neg_per_sim,
    sim_TP = actual_pos_per_sim - sim_FN,
    sim_TN = actual_neg_per_sim - sim_FP
  ) %>%
  group_by(Method) %>%
  summarize(
    sd_avg_TP = round(sd(sim_TP, na.rm = TRUE), 0),
    sd_avg_FP = round(sd(sim_FP, na.rm = TRUE), 0),
    sd_avg_FN = round(sd(sim_FN, na.rm = TRUE), 0),
    sd_avg_TN = round(sd(sim_TN, na.rm = TRUE), 0)
  )

cm_long_counts_sd <- cm_data_counts %>%
  pivot_longer(
    cols = c(avg_TP, avg_FP, avg_FN, avg_TN),
    names_to  = "Cell",
    values_to = "Count"
  ) %>%
  left_join(
    cm_sd_counts %>%
      pivot_longer(
        cols = c(sd_avg_TP, sd_avg_FP, sd_avg_FN, sd_avg_TN),
        names_to  = "Cell_sd",
        values_to = "SD"
      ) %>%
      mutate(Cell = sub("sd_avg_", "avg_", Cell_sd)) %>%
      dplyr::select(Method, Cell, SD),
    by = c("Method", "Cell")
  ) %>%
  mutate(
    Predicted = case_when(
      Cell %in% c("avg_TP", "avg_FP") ~ "Pred. Positive",
      TRUE                           ~ "Pred. Negative"
    ),
    Actual = case_when(
      Cell %in% c("avg_TP", "avg_FN") ~ "True Positive",
      TRUE                           ~ "True Negative"
    ),
    CellType = case_when(
      Cell == "avg_TP" ~ "TP",
      Cell == "avg_TN" ~ "TN",
      Cell == "avg_FP" ~ "FP",
      Cell == "avg_FN" ~ "FN"
    ),
    Label = paste0(Count, "±", SD)
  ) %>%
  mutate(
    Predicted = factor(Predicted, levels = c("Pred. Positive", "Pred. Negative")),
    Actual = factor(Actual, levels = c("True Positive", "True Negative"))
  )

cm_long_counts_sd$Method <- factor(cm_long_counts_sd$Method, levels = method_levels)

# ── Plot ──
ggplot(cm_long_counts_sd, aes(x = Predicted, y = Actual, fill = CellType)) +
  geom_tile(color = "white", linewidth = 0.8) +
  geom_text(aes(label = Label), size = 4, fontface = "bold", color = "white", lineheight = 0.9) +
  scale_fill_manual(
    values = c(
      "TP" = "#2E86AB",
      "TN" = "#2E86AB",
      "FN" = "#E84855",
      "FP" = "#F4A261"
    ),
    guide = guide_legend(title = "Cell type")
  ) +
  facet_wrap(~ Method, ncol = 3) +
  labs(
    title    = "Average Confusion Matrix per Simulation (± SD)",
    subtitle = paste0("Values are average counts ± SD out of 80 compounds | AET = ", AET, " µg/mL"),
    x = "Predicted class (reported concentration vs AET)",
    y = "Actual class (true concentration vs AET)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    strip.text       = element_text(face = "bold", size = 12),
    axis.text        = element_text(size = 9),
    panel.grid       = element_blank(),
    legend.position  = "bottom",
    plot.title       = element_text(face = "bold", size = 13),
    plot.subtitle    = element_text(size = 10, color = "grey40")
  )
```
